### 2022-12-12 draft1_Rcode.Rmd
Starting on creating code

---
title: "Untitled"
output: html_document
date: "2022-12-12"
---
```{r}
sample.dexp <- function(n,a,b,s,lamda){
  #Sampling from the double exp function g(x)
  #Using inverse transformation technique
  X <- NULL
  for (i in 1:n) {
      u <- runif(1) 
      if (u<0.5){
        X[i] <- b+log(2*u)/lamda
      }else{
        X[i] <- b-log(2-2*u)/lamda
      }
  }

  return(X)
}
dexp.sample <- sample.dexp(10000,2,3,1.25,0.5)
plot(density(dexp.sample))
#Plot looks good (double exp function with mean centered at 3)
#Note that its center (mean) is b and variance or spread is controlled by lambda (b on wiki note lambda=1/b) thus for smaller lambda variance increases
```
```{r}
q1.ar <- function(n,a,b,s,lambda,M){

    X <- NULL
    for (i in 1:n){
      repeat{
        #Drawing x from the propsal distibution 
        u <- runif(1)
        if (u<0.5){
          x <- b+log(2*u)/lambda
        }else {
          x <- b-log(2-2*u)/lambda
        }
        #Performing A-R sampling
        y <- runif(1)
        if (s < x){
          if(y <= (a*s^a)/(2*x^(a+1))/(M*lambda*exp(-lambda*abs(x-b))/2)){
            X[i]<-x
            break
          }
        } else if(s-b<x){
          if(y <= 1/(2*b)/(M*lambda*exp(-lambda*abs(x-b))/2)){
            X[i]<-x
            break
          }
        }
      }
    }
    return(X)
}

set.seed(2022)
ar.sample <- q1.ar(5000,2,3,1.25,0.5,M=3.2*exp(0.875))
#Creating function to create actual pdf of target distribution
target.dist <- function(x,a,b,s){
  n<-length(x)
  y<-rep(0,n)
  for (i in 1:n){

  if (x[i]<s-b){
    y[i] <- 0
  }else if (x[i]>s){
    y[i] <- (a*s^a)/(2*x[i]^(a+1))
  } else {
    y[i] <- 1/(2*b)
  }
  }
  return(y)
}

#Creating curve of target distribution
hist(ar.sample,nclass=100,prob = T,
     xlab = "Values generated by AR  sampling",
     main = "Simulated and actual pdf of target distribution");
curve(target.dist(x,2,3,1.25),from=-4,to=20,n=1000,add=T, lwd= 1.5, lty= 3,col="red");
legend("topright", lty = c(1, 3), cex = 0.6, col = c("black","red"),
legend = c("Simulated pdf", "Actual pdf"))
```

```{r}
q1.ar <- function(n,a,b,s,lambda,M){
  #Drawing x from the propsal distibution 
    X <- NULL
  for (i in 1:n){
    repeat{
    u <- runif(1)
    if (u<=exp(lambda*b)/2){
      x <- b +log(2*u)/lambda
  } else if (u>0.5){
    x <- b-log(2-2*u)/lambda
  }else {
    x <- 0
  }
    y <- runif(1)
    if (y <= M){
      X[i] <- x
      break
    }
    }
  }
    return(X)
}

set.seed(2022)
ar.sample <- q1.ar(5000,2,3,1.25,0.5,4)

hist(ar.sample,nclass=50,prob = T,
     xlab = "Values generated by AR  sampling",
     main = "Simulated and actual pdf of target distribution")
curve()

plot(density(ar.sample),lwd = 2,
main = "Simulated and actual pdf of Target distribution");

sum(ar.sample<(-1.75))
```

```{r}
q1.ar <- function(n,a,b,s,lambda,M){
  #Drawing x from the propsal distibution 
    X <- NULL
  for (i in 1:n){
    repeat{
    u <- runif(1)
    if (u<=0.5){
      x <- b +log(2*u)/lambda
  } else{
    x <- b-log(2-2*u)/lambda
  }
    y <- runif(1)
    if (y <= M){
      X[i] <- x
      break
    }
    }
  }
    return(X)
}

set.seed(2022)
ar.sample <- q1.ar(5000,2,3,1.25,0.5,4)

plot(density(ar.sample),lwd = 2,
main = "Simulated and actual pdf of Target distribution");

sum(ar.sample<(-1.75))
```
```{r}
#Creating actual pdf of pi(x)

target.dist <- function(x,a,b,s){

  n<-length(x)
  y<-rep(0,n)
  for (i in 1:n){

  if (x[i]<s-b){
    y[i] <- 0
  }else if (x[i]>s){
    y[i] <- (a*s^a)/(2*x[i]^(a+1))
  } else {
    y[i] <- 1/(2*b)
  }
  }
  return(y)
}

#Creating curve of target distribution
curve(target.dist(x,2,3,1.25),from=-4,to=10,n=1000)
```

### 2022-12-12 draft1_Rcode.Rmd
Getting rid of extra code when was figuring out question

@@
 ```
-
-```{r}
-q1.ar <- function(n,a,b,s,lambda,M){
-  #Drawing x from the propsal distibution 
-    X <- NULL
-  for (i in 1:n){
-    repeat{
-    u <- runif(1)
-    if (u<=exp(lambda*b)/2){
-      x <- b +log(2*u)/lambda
-  } else if (u>0.5){
-    x <- b-log(2-2*u)/lambda
-  }else {
-    x <- 0
-  }
-    y <- runif(1)
-    if (y <= M){
-      X[i] <- x
-      break
-    }
-    }
-  }
-    return(X)
-}
-
-set.seed(2022)
-ar.sample <- q1.ar(5000,2,3,1.25,0.5,4)
-
-hist(ar.sample,nclass=50,prob = T,
-     xlab = "Values generated by AR  sampling",
-     main = "Simulated and actual pdf of target distribution")
-curve()
-
-plot(density(ar.sample),lwd = 2,
-main = "Simulated and actual pdf of Target distribution");
-
-sum(ar.sample<(-1.75))
-```
-
-```{r}
-q1.ar <- function(n,a,b,s,lambda,M){
-  #Drawing x from the propsal distibution 
-    X <- NULL
-  for (i in 1:n){
-    repeat{
-    u <- runif(1)
-    if (u<=0.5){
-      x <- b +log(2*u)/lambda
-  } else{
-    x <- b-log(2-2*u)/lambda
-  }
-    y <- runif(1)
-    if (y <= M){
-      X[i] <- x
-      break
-    }
-    }
-  }
-    return(X)
-}
-
-set.seed(2022)
-ar.sample <- q1.ar(5000,2,3,1.25,0.5,4)
-
-plot(density(ar.sample),lwd = 2,
-main = "Simulated and actual pdf of Target distribution");
-
-sum(ar.sample<(-1.75))
-```
-```{r}
-#Creating actual pdf of pi(x)
-
-target.dist <- function(x,a,b,s){
-
-  n<-length(x)
-  y<-rep(0,n)
-  for (i in 1:n){
-
-  if (x[i]<s-b){
-    y[i] <- 0
-  }else if (x[i]>s){
-    y[i] <- (a*s^a)/(2*x[i]^(a+1))
-  } else {
-    y[i] <- 1/(2*b)
-  }
-  }
-  return(y)
-}
-
-#Creating curve of target distribution
-curve(target.dist(x,2,3,1.25),from=-4,to=10,n=1000)
-```

### 2022-12-12 draft1_Rcode.Rmd
Strugging to figure out Q2

@@
 ```{r}
+setwd("~/Desktop/Statistical Computing/CW/SC-CW2")
+```
+
+```{r}
+#Extra code to check was doing inverse transform sampling correctly
+
 sample.dexp <- function(n,a,b,s,lamda){
@@
 ```{r}
+## Question 1
+
+#c
 q1.ar <- function(n,a,b,s,lambda,M){
@@
 
+#d
 set.seed(2022)
@@
 legend = c("Simulated pdf", "Actual pdf"))
+
+#e
+#Using the plain Monte Carlo estimator of I (I_{n})
+
+#Function to calculate the kth root of kth moment using plain Monte Carlo estimator
+k.root.k.moment <- function(sample,n,k){
+  output <- (1/n*sum(sample^k))^(1/k)
+  return(output)
+}
+
+#k=2
+round(k.root.k.moment(ar.sample,5000,2),3)
+#2nd root of 2nd moment is 2.204 (3 d.p.)
+
+#k=3
+round(k.root.k.moment(ar.sample,5000,3),3)
+#3rd root of 3rd moment is 3.350 (3 d.p.)
 ```
+
+```{r}
+## Question 2
+n <- 100000
+X <- rexp(n,rate=0.5)+30 ## generate samples from shifted Exp(10/7)
+v.weights <- dnorm(X,mean=30,sd=0.7)/dexp(X-30,rate=0.5) ## vector of weights
+p.t <- mean(v.weights) ## calculate the IS estimator
+
+q2.is <- function(n,x0,time,a,lambda){
+
+}
+```
+
+```{r}
+## Question 5
+#a
+bootFStat <- function(group,y,w,bootCount){
+  #iterating for each replicate
+  for (b in 1:bootCount){
+    #creating bth bootstrap dataset
+    sample(y)
+  }
+
+}
+
+#b
+eco.df <- read.table("q5EcoStudy.txt",header = T)
+```

### 2022-12-13 draft1_Rcode.Rmd
Creating code for Q5 (check with code where know there are 3 groups)

@@
 ```{r}
-## Question 5
+#Q5 but with unknown number of groups
+
 #a
 bootFStat <- function(group,y,w,bootCount){
-  #iterating for each replicate
+  #object to store B bootstrap estimates of f statistic
+  Fstat.boot.est <- NULL
+  #grouping values into dataset 
+  data <- data.frame("group"=group,"Obs"=y,"weight"=w)
+  #Adding id column so can sample by id to retain all info of row (obs and weight)
+  data$id <- seq(1,dim(data)[1])
+
   for (b in 1:bootCount){
-    #creating bth bootstrap dataset
-    sample(y)
+    #object to store value of y tidle * N 
+    y.tilde.N <- 0
+
+    #loop to do bootstrap method for each group
+      for (group_val in unique(data$group)){
+        #selecting subset id's of dataset that is current group
+        group_dataset <- data$id[data$group==group_val]
+        #sampling from group dataset with replacement
+        group_sample.id <- sample(group_dataset, size = length(group_dataset), replace = TRUE)
+
+          #Using sampled id to select obs and weight for each id sampled to create bootstrap dataset for each group
+          #Selecting obs and weights using id's sampled
+          sample.group.obs <- NULL
+          sample.group.weight <- NULL
+
+          for (i in 1:length(group_sample.id)){
+            sample.group.obs[i] <- data$Obs[data$id == group_sample.id[i]]
+            sample.group.weight[i] <- data$weight[data$id == group_sample.id[i]]
+          }
+          #Setting sampled observations and weights to be those of group they belong to before repeating loop with next group
+          assign(paste0("sample.",group_val,".obs"),sample.group.obs)
+          assign(paste0("sample.",group_val,".weight"),sample.group.weight)
+
+          #Calculating y tilde_{i} for current group
+          y.tilde.group <- sum(sample.group.obs*sample.group.weight)/length(sample.group.obs)
+          assign(paste0("y.tilde.",group_val),y.tilde.group)
+
+          #updating y tilde using current group
+          y.tilde.N <- y.tilde.N + sum(sample.group.obs*sample.group.weight)
+
+          #Calculating part of denominator of f-statistic corresponding to current group
+          f.denom.group.part <- sum((sample.group.obs*sample.group.weight-y.tilde.group)^2)
+          assign(paste0("f.denom.",group_val),f.denom.group.part)
       }
 
+    #calculating y tilde
+    y.tilde <- y.tilde.N/length(y)
+    # y.tilde <- y.tilde.N/300
+
+    #calculating summation parts of numerator and denominator of f statistic
+    fstat.num <- 0
+    fstat.denom <- 0
+    for (group_val in unique(data$group)){
+      fstat.num <- fstat.num+length(get(paste0("sample.",group_val,".obs")))*(get(paste0("y.tilde.",group_val))-y.tilde)^2
+      fstat.denom <- fstat.denom + get(paste0("f.denom.",group_val))
+    }
+    #calculating F stat
+    fstat <- (fstat.num/(length(unique(group))-1))/(fstat.denom/(length(y)-length(unique(group))))
+    Fstat.boot.est[b] <- fstat
+  }
+
+  #Calculating standard error of bootstrap f statistic
+  bootstr.mu <- sum(Fstat.boot.est)/length(Fstat.boot.est)
+  bootstr.se <- sqrt(sum((Fstat.boot.est-bootstr.mu )^2)/(length(Fstat.boot.est-1)))
+
+  return(bootstr.se)
 }
@@
 #b
+set.seed(2022)
+bootFStat(eco.df$habitat,eco.df$density,eco.df$weight,1000)
+
+```
+
+```{r}
+## Question 5 extra stuff
+
+#a
+unique(eco.df$habitat)[3]
+
+bootFStat <- function(group,y,w,bootCount){
+  #grouping into dataset 
+  data <- data.frame("group"=group,"Obs"=y,"weight"=w)
+  #Adding id column so can sample by id to retain all info of row (obs and weight)
+  data$id <- seq(1,dim(data)[1])
+
+  #Splitting data into groups depending on group
+  group.A <- data$id[data$group=="A"]
+  group.B <- data$id[data$group=="B"]
+  group.C <- data$id[data$group=="C"]
+  #Sampling with replacement
+  sample.A.id <- sample(group.A, size = length(group.A), replace = TRUE)
+  sample.B.id <- sample(group.B, size = length(group.B), replace = TRUE)
+  sample.C.id <- sample(group.C, size = length(group.C), replace = TRUE)
+
+  #Using sampled id to select obs and weight for each id sampled to create bootstrap dataset for each group
+  #Selecting obs and weights using id's sampled
+
+  #For A
+  sample.A.obs <- NULL
+  sample.A.weight <- NULL
+  for (i in 1:length(sample.A.id)){
+    sample.A.obs[i] <- data$Obs[data$id == sample.A.id[i]]
+    sample.A.weight[i] <- data$weight[data$id == sample.A.id[i]]
+  }
+  #calculating y.tilde_{i} for A
+  y.tilde.A <- sum(sample.A.obs*sample.A.obs)/length(sample.A.obs)
+
+  #For B
+  sample.B.obs <- NULL
+  sample.B.weight <- NULL
+  for (i in 1:length(sample.B.id)){
+    sample.B.obs[i] <- data$Obs[data$id == sample.B.id[i]]
+    sample.B.weight[i] <- data$weight[data$id == sample.B.id[i]]
+  }
+  #calculating y.tilde_{i} for B
+  y.tilde.B <- sum(sample.B.obs*sample.B.obs)/length(sample.B.obs)
+
+  #For C
+  sample.C.obs <- NULL
+  sample.C.weight <- NULL
+  for (i in 1:length(sample.C.id)){
+    sample.C.obs[i] <- data$Obs[data$id == sample.C.id[i]]
+    sample.C.weight[i] <- data$weight[data$id == sample.C.id[i]]
+  }
+  #calculating y.tilde_{i} for C
+  y.tilde.C <- sum(sample.C.obs*sample.C.obs)/length(sample.C.obs)
+
+  #Calculating y.tilde
+  y.tilde <- (y.tilde.A*length(sample.A.obs)+y.tilde.B*length(sample.B.obs)+y.tilde.C*length(sample.C.obs))/length(y)
+
+  #Calculating F.tilde
+  F.num <- 1
+  F.denom <-2
+}
+
+library(dplyr)
 eco.df <- read.table("q5EcoStudy.txt",header = T)
+data <- eco.df%>%
+  rename("group"="habitat","Obs"="density")
+data$id <- seq(1,dim(data)[1])
+
+  #Splitting data into groups depending on group
+  group.A <- data$id[data$group=="A"]
+  group.B <- data$id[data$group=="B"]
+  group.C <- data$id[data$group=="C"]
+  #Sampling with replacement
+  sample.A.id <- sample(group.A, size = length(group.A), replace = TRUE)
+  sample.B.id <- sample(group.B, size = length(group.B), replace = TRUE)
+  sample.C.id <- sample(group.C, size = length(group.C), replace = TRUE)
+
+  #Using sampled id to select obs and weight for each id sampled to create bootstrap dataset for each group
+  #Selecting obs and weights using id's sampled
+
+  #For A
+  sample.A.obs <- NULL
+  sample.A.weight <- NULL
+  for (i in 1:length(sample.A.id)){
+    sample.A.obs[i] <- data$Obs[data$id == sample.A.id[i]]
+    sample.A.weight[i] <- data$weight[data$id == sample.A.id[i]]
+  }
+  #calculating y.tilde_{i} for A
+  y.tilde.A <- sum(sample.A.obs*sample.A.obs)/length(sample.A.obs)
+
+#b
+eco.df <- read.table("q5EcoStudy.txt",header = T)
+
+test
 ```

### 2022-12-13 draft1_Rcode.Rmd
Updating Q5 so that sampling does not have to select a certain number from each group

@@
 
+#Need to edit as in bootstrapping dont care aout having same number in each group! (I think)
+
 #a
@@
 ```{r}
-## Question 5 extra stuff
+#Bootstrap where number in each group not set
+
+bootFStat <- function(group,y,w,bootCount){
+  #object to store B bootstrap estimates of f statistic
+  Fstat.boot.est <- NULL
+  #grouping values into dataset 
+  data <- data.frame("group"=group,"Obs"=y,"weight"=w)
+  #Adding id column so can sample by id to retain all info of row (obs and weight)
+  data$id <- seq(1,dim(data)[1])
+
+  for (b in 1:bootCount){
+    #object to store value of y tilde * N 
+    y.tilde.N <- 0
+
+    #sampling dataset
+    sample.id <- sample(data$id, size = length(y), replace = TRUE)
+    #extracting weight and y value corresponding to id sampled
+    sample.obs <- NULL
+    sample.weight <- NULL
+    sample.group <- NULL
+
+    for (i in 1:length(y)){
+      sample.obs[i] <- data$Obs[data$id == sample.id[i]]
+      sample.weight[i] <- data$weight[data$id == sample.id[i]]
+      sample.group[i] <- data$group[data$id == sample.id[i]]
+    }
+    #creating dataset of sample
+    sample.dataset <- data.frame("group"=sample.group,"Obs"=sample.obs,"weight"=sample.weight)
+
+    #computing statistics to calculate f-stat
+    for (group_val in unique(sample.dataset$group)){
+      group.dataset <- data.frame("Obs"=sample.dataset$Obs[sample.dataset==group_val],
+                                  "weight"=sample.dataset$weight[sample.dataset==group_val])
+
+      #calculating number of obs for current group
+      assign(paste0("nobs.",group_val),dim(group.dataset)[1])
+
+      #Calculating y tilde_{i} for current group
+      y.tilde.group <- sum(group.dataset$Obs*group.dataset$weight)/dim(group.dataset)[1]
+      assign(paste0("y.tilde.",group_val),y.tilde.group)
+
+      #updating y tilde using current group
+      y.tilde.N <- y.tilde.N + sum(group.dataset$Obs*group.dataset$weight)
+
+      #Calculating part of denominator of f-statistic corresponding to current group
+      f.denom.group.part <- sum((group.dataset$Obs*group.dataset$weight-y.tilde.group)^2)
+      assign(paste0("f.denom.",group_val),f.denom.group.part)
+
+    }
+    #calculating y tilde
+    y.tilde <- y.tilde.N/length(y)
+
+    #calculating summation parts of numerator and denominator of f statistic
+    fstat.num <- 0
+    fstat.denom <- 0
+    for (group_val in unique(data$group)){
+      fstat.num <- fstat.num+get(paste0("nobs.",group_val))*(get(paste0("y.tilde.",group_val))-y.tilde)^2
+      fstat.denom <- fstat.denom + get(paste0("f.denom.",group_val))
+    }
+    #calculating F stat
+    fstat <- (fstat.num/(length(unique(group))-1))/(fstat.denom/(length(y)-length(unique(group))))
+    Fstat.boot.est[b] <- fstat
+
+  }
+  #Calculating standard error of bootstrap f statistic
+  bootstr.mu <- sum(Fstat.boot.est)/length(Fstat.boot.est)
+  bootstr.se <- sqrt(sum((Fstat.boot.est-bootstr.mu )^2)/(length(Fstat.boot.est-1)))
+
+  return(bootstr.se)
+}
+
+#b
+set.seed(2022)
+bootFStat(eco.df$habitat,eco.df$density,eco.df$weight,1000)  
+```
+
+```{r}
+## Question 5 extra stuff (use this to check results by doing piece by piece)
 

### 2022-12-13 draft1_Rcode.Rmd
Working on Question 4

@@
 ```{r}
+## Question 4
+q4.data <- read.csv("q4MixLinReg.csv",header=T)
+
+#Part a
+linRegMixEstep <- function(piCur,aCur,bCur,sCur,x,Y){
+  #matrix to store E-Step values
+  #row i (p(Z_{i} = 1|Y_{i}, θ^{cur}),..., p(Z_{i} = K|Y_{i}, θ^{cur})) row i has X_{i} ie X fixed
+  #column j (p(Z_{i} = j|Y_{1}, θ^{cur}), ...,(Z_{i} = k|Y_{k}, θ^{cur})) column j has Z_{i}=j  ie value of Z_{i} fixed
+  prob <- matrix(data=NA,nrow = length(x),ncol = length(piCur))
+  for (i in 1:length(x)){
+    for (j in 1:length(piCur)){
+      prob[i,j] <- piCur[j]*dnorm(Y[i],mean=aCur[j]+bCur[j]*x[i],sd=sCur[j])/sum(piCur*dnorm(Y[i],mean=aCur+bCur*x[i],sd=sCur))
+    }
+  }
+  return(prob)
+}
+
+#test <-linRegMixEstep(piCur=c(0.4,0.3,0.3),aCur=c(0.1,-0.1,0.2),bCur=c(1,-1,1),sCur=c(1,0.5,1.1),x=q4.data$x,Y=q4.data$Y)
+
+#Part b
+#i
+calcNewCoefs <- function(W,x,Y){
+  #Calculating a^{new) and b^{new}
+  new.coef <- matrix(data=NA,ncol = 2,nrow=dim(W)[2])
+  #Creating design matrix
+  X <- matrix(c(rep(1,length(x)),x),ncol=2)
+  for (k in 1:dim(W)[2]){
+    new.coef[k,] <- solve(t(X)%*%diag(W[,k])%*%X)%*%t(X)%*%diag(W[,k])%*%Y
+  }
+  return(t(new.coef))
+}
+#new.coef <- calcNewCoefs(test,x=q4.data$x,Y=q4.data$Y)
+
+#ii
+calcNewSd <- function(W,x,Y,a,b){
+  new.sd <- NULL
+  #calculating new sd
+  for (k in 1:dim(W)[2]){
+    new.sd[k] <- sqrt(sum(W[,k]*(Y-a[k]-b[k]*x)^2)/sum(W[,k]))
+  }
+  return(new.sd)
+}
+#new.sd <- calcNewSd(test,x,Y,new.coef[1,],new.coef[2,])
+
+#iii
+linRegMixMstep <- function(W,x,Y){
+  #calculating new coefficents for a and b
+  new.coef <- calcNewCoefs(W,x,Y)
+
+  #calculating new sd
+  new.sd <- calcNewSd(W,x,Y,new.coef[1,],new.coef[2,])
+
+  #Calculating new pi
+  new.pi <- NULL
+  for (k in 1:dim(W)[2]){
+    new.pi[k] <- sum(W[,k])/dim(W)[1]
+  }
+ return(list(new.pi,new.coef[1,],new.coef[2,],new.sd))
+}
+#new.theta <- linRegMixMstep(test,x,Y)
+
+#c
+###############This is wrong!!!!! Need to change (covergence isn't working, plan to use final values from final function in d to see what is happening)
+linRegMixCalcLogLik <- function(x,Y,piCur,piNew,aCur,aNew,bCur,bNew,sCur,sNew){
+  #calculating conditional probabilities for current and new values of theta
+  W.cur <- linRegMixEstep(piCur,aCur,bCur,sCur,x,Y)
+  W.new <- linRegMixEstep(piNew,aNew,bNew,sNew,x,Y)
+  #Creating object to store values of log-likelihood for each observation i of Y 
+  Q.cur <- NULL
+  Q.new <- NULL
+  for (i in 1:length(Y)){
+    #calculating log likelihood for each value of Y
+    Q.cur[i] <- sum(W.cur[i,]*log(piCur))+sum(W.cur[i,]*log(dnorm(Y[i],mean = aCur+bCur*x[i],sd=sCur)))
+    Q.new[i] <- sum(W.new[i,]*log(piNew))+sum(W.new[i,]*log(dnorm(Y[i],mean = aNew+bNew*x[i],sd=sNew)))
+  }
+  #Summing over all observations as is in the formula
+  Q.cur <- sum(Q.cur)
+  Q.new <- sum(Q.new)
+  return(c(Q.cur,Q.new))
+}
+
+#linRegMixCalcLogLik(x,Y,piCur,new.pi,aCur,new.coef[1,],bCur,new.coef[2,],sCur,new.sd)
+
+#d
+linRegMixEM <- function(piInit,aInit,bInit,sInit,x,Y,convergeEps){
+  #Initializing the algorithm
+  piCur <- piInit
+  aCur <- aInit
+  bCur <- bInit
+  sCur <- sInit
+
+  conv.critera=0
+  i=0
+  while(i==0){
+    #E-step
+    W <- linRegMixEstep(piCur,aCur,bCur,sCur,x,Y)
+    #Calculating new values for parameters
+    theta.new <- linRegMixMstep(W,x,Y)
+    #Calculating log-likelihood for current and new values of theta
+    theta.loglik <- linRegMixCalcLogLik(x,Y,piCur,theta.new[[1]],aCur,theta.new[[2]],bCur,theta.new[[3]],sCur,theta.new[[4]])
+    if(abs(theta.loglik[2]-theta.loglik[1])<convergeEps){
+      conv.critera=1
+    }
+
+  }
+  return(theta.new )
+}
+
+test <- linRegMixEM(c(0.4,0.3,0.3),c(0.1,-0.1,0.2),c(1,-1,1),c(1,0.5,1.1),q4.data$x,q4.data$Y,1e-06)
+
+```
+
+```{r}
+linRegMixEM <- function(piInit,aInit,bInit,sInit,x,Y,convergeEps,test){
+  #Initializing the algorithm
+  piCur <- piInit
+  aCur <- aInit
+  bCur <- bInit
+  sCur <- sInit
+
+  conv.critera=0
+
+  while(i<test){
+    #E-step
+    W <- linRegMixEstep(piCur,aCur,bCur,sCur,x,Y)
+    #Calculating new values for parameters
+    theta.new <- linRegMixMstep(W,x,Y)
+    #Calculating log-likelihood for current and new values of theta
+    theta.loglik <- linRegMixCalcLogLik(x,Y,piCur,theta.new[[1]],aCur,theta.new[[2]],bCur,theta.new[[3]],sCur,theta.new[[4]])
+
+    i=i+1
+
+  }
+  return(theta.new )
+}
+
+test.1010 <- linRegMixEM(c(0.4,0.3,0.3),c(0.1,-0.1,0.2),c(1,-1,1),c(1,0.5,1.1),q4.data$x,q4.data$Y,1e-06,100)
+```
+
+```{r}
 #Q5 but with unknown number of groups
@@
     sample.id <- sample(data$id, size = length(y), replace = TRUE)
-    #extracting weight and y value corresponding to id sampled
+
     sample.obs <- NULL
@@
     sample.group <- NULL
-
+    #extracting weight, group and y value corresponding to id sampled
     for (i in 1:length(y)){
@@
 }
+#Sorry about using so many loops :P
 

### 2022-12-14 draft1_Rcode.Rmd
Fixed question 4 (now converges to an answer)

--- 2022-12-13 draft1_Rcode.Rmd
@@
 ```{r}
+## Question 3
+
+```
+
+```{r}
+## Question 4
+q4.data <- read.csv("q4MixLinReg.csv",header=T)
+
+#Part a
+linRegMixEstep <- function(piCur,aCur,bCur,sCur,x,Y){
+  #matrix to store E-Step values
+  #row i (p(Z_{i} = 1|Y_{i}, θ^{cur}),..., p(Z_{i} = K|Y_{i}, θ^{cur})) row i has X_{i} ie X fixed
+  #column j (p(Z_{i} = j|Y_{1}, θ^{cur}), ...,(Z_{i} = k|Y_{k}, θ^{cur})) column j has Z_{i}=j  ie value of Z_{i} fixed
+  prob <- matrix(data=NA,nrow = length(x),ncol = length(piCur))
+  for (i in 1:length(x)){
+    for (j in 1:length(piCur)){
+      prob[i,j] <- piCur[j]*dnorm(Y[i],mean=aCur[j]+bCur[j]*x[i],sd=sCur[j])/sum(piCur*dnorm(Y[i],mean=aCur+bCur*x[i],sd=sCur))
+    }
+  }
+  return(prob)
+}
+
+#Part b
+#i
+calcNewCoefs <- function(W,x,Y){
+  #Calculating a^{new) and b^{new}
+  new.coef <- matrix(data=NA,ncol = 2,nrow=dim(W)[2])
+  #Creating design matrix
+  X <- matrix(c(rep(1,length(x)),x),ncol=2)
+  for (k in 1:dim(W)[2]){
+    new.coef[k,] <- solve(t(X)%*%diag(W[,k])%*%X)%*%t(X)%*%diag(W[,k])%*%Y
+  }
+  return(t(new.coef))
+}
+
+#ii
+calcNewSd <- function(W,x,Y,a,b){
+  new.sd <- NULL
+  #calculating new sd
+  for (k in 1:dim(W)[2]){
+    new.sd[k] <- sqrt(sum(W[,k]*(Y-a[k]-b[k]*x)^2)/sum(W[,k]))
+  }
+  return(new.sd)
+}
+
+#iii
+linRegMixMstep <- function(W,x,Y){
+  #calculating new coefficents for a and b
+  new.coef <- calcNewCoefs(W,x,Y)
+
+  #calculating new sd
+  new.sd <- calcNewSd(W,x,Y,new.coef[1,],new.coef[2,])
+
+  #Calculating new pi
+  new.pi <- NULL
+  for (k in 1:dim(W)[2]){
+    new.pi[k] <- sum(W[,k])/dim(W)[1]
+  }
+ return(list(new.pi,new.coef[1,],new.coef[2,],new.sd))
+}
+
+#Part c
+linRegMixCalcLogLik <- function(x,Y,piCur,piNew,aCur,aNew,bCur,bNew,sCur,sNew){
+  #calculating conditional probabilities for current and new values of theta
+  W.cur <- linRegMixEstep(piCur,aCur,bCur,sCur,x,Y)
+  W.new <- linRegMixEstep(piNew,aNew,bNew,sNew,x,Y)
+  #Creating object to store values of log-likelihood for each observation i of Y 
+  Q.cur <- NULL
+  Q.new <- NULL
+  for (i in 1:length(Y)){
+    #calculating log likelihood for each value of Y
+    Q.cur[i] <- sum(W.cur[i,]*log(piCur))+sum(W.cur[i,]*log(dnorm(Y[i],mean = aCur+bCur*x[i],sd=sCur)))
+    Q.new[i] <- sum(W.new[i,]*log(piNew))+sum(W.new[i,]*log(dnorm(Y[i],mean = aNew+bNew*x[i],sd=sNew)))
+  }
+  #Summing over all observations as is in the formula
+  Q.cur <- sum(Q.cur)
+  Q.new <- sum(Q.new)
+  return(c(Q.cur,Q.new))
+}
+
+#Part d
+linRegMixEM <- function(piInit,aInit,bInit,sInit,x,Y,convergeEps){
+  #Initializing the algorithm
+  piCur <- piInit
+  aCur <- aInit
+  bCur <- bInit
+  sCur <- sInit
+
+  conv.critera <- 0
+  while(conv.critera==0){
+    #E-step
+    W <- linRegMixEstep(piCur,aCur,bCur,sCur,x,Y)
+    #Calculating new values for parameters
+    theta.new <- linRegMixMstep(W,x,Y)
+    #Calculating log-likelihood for current and new values of parameters
+    theta.loglik <- linRegMixCalcLogLik(x,Y,piCur,theta.new[[1]],aCur,theta.new[[2]],bCur,theta.new[[3]],sCur,theta.new[[4]])
+    #checking convergence criteria
+    if(abs(theta.loglik[2]-theta.loglik[1])<convergeEps){
+      conv.critera=1
+    }
+    #Updating current values of parameters
+    piCur <- theta.new[[1]]
+    aCur <- theta.new[[2]]
+    bCur <- theta.new[[3]]
+    sCur <- theta.new[[4]]
+  }
+  return(theta.new)
+}
+
+#Part e
+mle.est <- linRegMixEM(c(0.4,0.3,0.3),c(0.1,-0.1,0.2),c(1,-1,1),c(1,0.5,1.1),q4.data$x,q4.data$Y,1e-06)
+#pi MLEs
+round(mle.est[[1]],3)
+# (0.450, 0.255, 0.295)
+
+#alpha MLEs
+round(mle.est[[2]],3)
+# (0.838,  7.458, -5.308)
+
+#beta MLEs
+round(mle.est[[3]],3)
+# (0.258, -1.755,  1.223)
+
+#sd MLEs
+round(mle.est[[4]],3)
+# (0.787, 0.935, 1.450)
+```
+
+```{r}
 #Q5 but with unknown number of groups

### 2022-12-15 Figuring_Q3.Rmd
Trying to figure out question 3

---
title: "Untitled"
output: html_document
date: "2022-12-14"
---
```{r}
# Part b
calcDirichletPDF <- function(x,a){
  #Calculating density of x under Dirichlet distribution with parameters specified by a
  density <- (gamma(sum(a))/prod(gamma(a)))*prod(x^(a-1))
  #calculating log of density
  log.density <- log(density)
  #returning a numeric vector
  return(c(log.density))
}

#Part c
simDirichletRV <- function(a){
  #Drawing from gamma distribution
  U <- rgamma(a,1)
  #Computing V a vector drawn from Dirichlet distribution with parameters specified by a
  V <- U/sum(U)

  return(V)
}

###Testing 
#Initial values
#finding prob of sequence under current pt
X <- read.table("q3Weather.txt",header = F)
X <- X$V1
initPT <- matrix(1/3,nrow=3,ncol=3)
initStateProbs <- c(1/3,1/3,1/3)
n <- 100
burnin <- 5

  #Converting X into a numeric vector to access entries of PT
  X <- ifelse(X=="s",1,(ifelse(X=="c",2,3)))
  init.state <- X[1:(length(X)-1)]
  trans.state <- X[2:length(X)]

  #initializing algorithm
  V <- array(NA,dim=c(3,3,n+burnin+1))
  V[,,1] <- initStateProbs

proposal <- matrix(NA,ncol=3,nrow=3)
#starting loop
j=2
    #Generating proposal value for each row of transition matrix
    for (i in 1:length(initStateProbs)){
      proposal[i,] <- simDirichletRV(20*V[i,,j-1])
    }
     #calculating alpha

      #Generating posterior using equation (12) for proposal and old value of PT
      actualp.prev <- actualp.prop <- NULL
      for (i in 1:length(init.state)){
        actualp.prev[i] <- V[init.state[i],trans.state[i],j-1]
        actualp.prop[i] <- proposal[init.state[i],trans.state[i]]
      }
      #density has some constant as proportional but due to fraction constant cancels out so dont
      #need to worry about it
      true.density.prev <- prod(actualp.prev)*initStateProbs[X[1]]
      true.density.prop <-prod(actualp.prop)*initStateProbs[X[1]]
      true.density.prop/true.density.prev

      #Doing shady stuff (assuming that px` or previous is the x vector :P)
      calcDirichletPDF(x=proposal[2,],a=V[2,,j-1])

# A <- array(NA,dim=c(3,3,n+burnin+1))
# A[,,1] <- matrix(c(1,2,3,4,5,6,7,8,9),ncol=3,nrow=3,byrow = T)
# #1st index is row
# A[1,,1]
# #2nd index is column
# A[,1,1]
```

```{r}
## Question 3

#start by doing b and c 
#then do d awhile figuring out a (the algorithm part)

#Part b
calcDirichletPDF <- function(x,a){
  #Calculating density of x under Dirichlet distribution with parameters specified by a
  density <- (gamma(sum(a))/prod(gamma(a)))*prod(x^(a-1))
  #calculating log of density
  log.density <- log(density)
  #returning a numeric vector
  return(c(log.density))
}

#Part c
simDirichletRV <- function(a){
  #Drawing from gamma distribution
  U <- rgamma(a,1)
  #Computing V a vector drawn from Dirichlet distribution with parameters specified by a
  V <- U/sum(U)

  return(V)
}

#Part d
q3.mcmc <- function(X,initStateProbs,initPT,burnin,n){
  #X is vector of weather for each observed day (d long if recorded weather for d days)
  #Converting X into a numeric vector to access entries of PT
  X <- ifelse(X=="s",1,(ifelse(X=="c",2,3)))
  init.state <- X[1:(length(X)-1)]
  trans.state <- X[2:length(X)]

  #initializing algorithm
  V <- array(NA,dim=c(3,3,n+burnin+1))
  V[,,1] <- initStateProbs

  for (j in 2:(n+burnin+1)){
    proposal <- matrix(NA,ncol=3,nrow=3)
    #Generating proposal value for each row of transition matrix
    for (i in 1:length(initPT)){
      proposal[i,] <- simDirichletRV(20*V[,i,j-1])
    }

      #Generating posterior using equation (12) for proposal and old value of PT
      prob.trans.prev <- NULL
      prob.trans.curr <- NULL
      for (i in 1:length(init.state)){
        prob.trans.prev[i] <- V[init.state[i],trans.state[i],j-1]
        prob.trans.curr[i] <- proposal[trans.state[i],init.state[i]]
      }
      #density has some constant as proportional but due to fraction constant cancels out so dont
      #need to worry about it
      density.prev <- prod(prob.trans.prev)*initStateProbs[X[1]]
      density.curr <-prod(prob.trans.curr)*initStateProbs[X[1]]

      #alpha simplier as q is symmetric (Big boi assumption may not be able to do this)
      alpha <- min(1,density.curr/density.prev)

      #Accept or Reject
      if (runif(1) <= alpha){
        V[,,j] <- proposal
      } else{
        V[,,j] <- V[,,j-1]
      }
  }

  return(V[,,(burnin+1:n+burnin+1)])

}
#finding prob of sequence under current pt
X <- read.table("q3Weather.txt",header = F)
X <- X$V1

#Part e
initPT <- matrix(1/3,nrow=3,ncol=3)
initStateProbs <- c(1/3,1/3,1/3)

q3.mcmc(X,initStateProbs,initPT,10000,100000)
```

### 2022-12-15 draft1_Rcode.Rmd
Trying to figure out question 3

@@
 
+#start by doing b and c 
+#then do d awhile figuring out a (the algorithm part)
+
+#Part b
+calcDirichletPDF <- function(x,a){
+  #Calculating density of x under Dirichlet distribution with parameters specified by a
+  density <- (gamma(sum(a))/prod(gamma(a)))*prod(x^(a-1))
+  #calculating log of density
+  log.density <- log(density)
+  #returning a numeric vector
+  return(c(log.density))
+}
+
+#Part c
+simDirichletRV <- function(a){
+  #Drawing from gamma distribution
+  U <- rgamma(a,1)
+  #Computing V a vector drawn from Dirichlet distribution with parameters specified by a
+  V <- U/sum(U)
+
+  return(V)
+}
+
+#Part d
+q3.mcmc <- function(X,initStateProbs,initPT,burnin,n){
+  #X is vector of weather for each observed day (d long if recorded weather for d days)
+  #Converting X into a numeric vector to access entries of PT
+  X <- ifelse(X=="s",1,(ifelse(X=="c",2,3)))
+  init.state <- X[1:(length(X)-1)]
+  trans.state <- X[2:length(X)]
+
+  #initializing algorithm
+  V <- array(NA,dim=c(3,3,n+burnin+1))
+  V[,,1] <- initStateProbs
+
+  for (j in 2:(n+burnin+1)){
+    proposal <- matrix(NA,ncol=3,nrow=3)
+    #Generating proposal value for each row of transition matrix
+    for (i in 1:length(initPT)){
+      proposal[i,] <- simDirichletRV(20*V[,i,j-1])
+    }
+
+      #Generating posterior using equation (12) for proposal and old value of PT
+      prob.trans.prev <- NULL
+      prob.trans.curr <- NULL
+      for (i in 1:length(init.state)){
+        prob.trans.prev[i] <- V[init.state[i],trans.state[i],j-1]
+        prob.trans.curr[i] <- proposal[trans.state[i],init.state[i]]
+      }
+      #density has some constant as proportional but due to fraction constant cancels out so dont
+      #need to worry about it
+      density.prev <- prod(prob.trans.prev)*initStateProbs[X[1]]
+      density.curr <-prod(prob.trans.curr)*initStateProbs[X[1]]
+
+      #alpha simplier as q is symmetric (Big boi assumption may not be able to do this)
+      alpha <- min(1,density.curr/density.prev)
+
+      #Accept or Reject
+      if (runif(1) <= alpha){
+        V[,,j] <- proposal
+      } else{
+        V[,,j] <- V[,,j-1]
+      }
+  }
+
+  return(V[,,(burnin+1:n+burnin+1)])
+
+}
+#finding prob of sequence under current pt
+X <- read.table("q3Weather.txt",header = F)
+X <- X$V1
+
+#Part e
+initPT <- matrix(1/3,nrow=3,ncol=3)
+initStateProbs <- c(1/3,1/3,1/3)
+
+q3.mcmc(X,initStateProbs,initPT,10000,100000)
+```
+
+```{r}
+  #initializing algorithm
+  V <- array(NA,dim=c(3,3,500))
+  V[,,1] <- initStateProbs
+
+  V
+
+  for (j in 2:(n+burnin+1)){
+    proposal <- matrix(NA,ncol=3,nrow=3)
+    #Generating proposal value for each row of transition matrix
+    for (i in 1:length(initPT)){
+      x <- V[,i,j-1]
+      proposal[1,] <- simDirichletRV(20*V[,1,2-1])
+    }
+   test <- matrix(c(1,2,3,4),ncol=2,nrow=2) 
+test[2,1]
 ```

### 2022-12-17 Figuring_Q3.Rmd
Starting to work on code producing pdf report

@@
 initStateProbs <- c(1/3,1/3,1/3)
-n <- 100
+n <- 500000
 burnin <- 5
@@
   V[,,1] <- initStateProbs
+  est.state.prob <- array(NA,dim=c(1,3,n+burnin+1)) 
+  est.state.prob[,,1] <-  initStateProbs
 
@@
 #starting loop
-j=2
+for (j in 2:(n+burnin+1)){
+
+    est.state.prob[,,j] <- est.state.prob[,,j-1]%*%V[,,j-1]
     #Generating proposal value for each row of transition matrix
@@
       true.density.prop <-prod(actualp.prop)*initStateProbs[X[1]]
-      true.density.prop/true.density.prev
+      #alpha <- true.density.prop/true.density.prev
 
       #Doing shady stuff (assuming that px` or previous is the x vector :P)
-      calcDirichletPDF(x=proposal[2,],a=V[2,,j-1])
+      #Calculating pdf under proposal using x as the probability to be in each state 
+       prop.density.prop <- exp(calcDirichletPDF(x=proposal[2,],a=est.state.prob[,,j-1])+
+                                  calcDirichletPDF(x=proposal[1,],a=est.state.prob[,,j-1])+calcDirichletPDF(x=proposal[3,],a=est.state.prob[,,j-1]))
+      # #same but under previous accepted value
+       prop.density.prev <- exp(calcDirichletPDF(x=V[2,,j-1],a=est.state.prob[,,j-1])+
+                                 calcDirichletPDF(x=V[1,,j-1],a=est.state.prob[,,j-1])+calcDirichletPDF(x=V[3,,j-1],a=est.state.prob[,,j-1]))
+
+      # alpha <- (true.density.prop*prop.density.prev)/(true.density.prev*prop.density.prop)
+      alpha <-  prop.density.prop/prop.density.prev
+
+      if(min(1,alpha)>=runif(1)){
+        V[,,j] <- proposal
+      }else{
+        V[,,j] <- V[,,j-1]
+      }
+}
 
@@
 # A[,1,1]
+
+plot(V[1,2,],type="l") #No convegerence :(
 ```

### 2022-12-17 draft_report.Rmd
Starting to work on code producing pdf report

---
title: "STatistical Computing (MATH6173) Coursework 2"
author: "Student ID: 34273638"
geometry: margin = 2cm
output:
  pdf_document: default
header-includes:
  - \usepackage{wrapfig}
  - \usepackage{lipsum}
  - \usepackage{setspace}
  - \usepackage{titlesec}
  - \usepackage{amsmath}
  - \usepackage{amssymb}
  - \usepackage{mathtools}
  - \titlespacing{\title}{0pt}{\parskip}{-\parskip}
---

```{r, include=F}
#Setting up enviroment
setwd("~/Desktop/Statistical Computing/CW/SC-CW2")
```

# Question 1

## Part a: (add graphs to help in explanation)

\begin{equation}
\pi(x) =
\begin{cases}
 \frac{as^{a}}{2x^{a+1}}\, ,& s<x, \\
 \frac{1}{2b}\, ,& s-b \leq x \leq s, \\
 0  ,& \text{otherwise,}
 \end{cases}  
 \end{equation}

\begin{equation}
g(x) = \frac{\lambda}{2}exp(-\lambda|x-b|)
\end{equation}

For acceptance-rejection sampling we need to find constant $M>0$ such that $\pi(x) \leq Mg(x)$ for $\pi(x)$ the target distribution and $g(x)$ the proposal density. This constant M determines the acceptance probability and hence the efficiency of the algorithm. With the acceptance probability (denoted $\tau$) equal to $\frac{1}{M}$. For multiple possible values of $\lambda$ in $g(x)$ the value of $\lambda$ that results in a smaller constant M will have the higher acceptance probability. To find M for each $\lambda$ value I split the real domain of x into 3 intervals corresponding to the different PARTS of $\pi$ and study the relationship between $g(x)$ and $\pi(x)$ within these 3 intervals to find $M_{1},M_{2},M_{3}$ which correspond to the smallest constant such that $\pi(x) \leq M_{i}g(x)$ for the ith interval. Then $M=max(M_{1},M_{2},M_{3})$ as this is the smallest constant that satisfies $\pi(x) \leq Mg(x)$ for all possible values of x.

For the first region ($-\infty, s-b$) $\pi(x)=0$ thus we can let $M_{1}$ be some arbitrarily small value greater than 0 say $\epsilon$ and this will satisfy $\pi(x) \leq M_{1}g(x)$ in the region specified.
For the second region [$s-b,s$], $\pi(x)=\frac{1}{2b}$. Considering that $g(x)$ is an increasing function in this region it will be at it's minimum value at $x=s-b$ and that the value of $\pi(x)$ is constant, setting $M_{2}=\frac{\pi(s-b)}{g(s-b)}=\frac{exp(\lambda|s-2b|)}{\lambda b}$ will satisfy the equation $\pi(x) \leq M_{2}g(x)$ for all x in the region.
For the third region ($s,\infty$), $\pi(x)=\frac{as^{a}}{2x^{a+1}}$. I note that $\pi(x)$ is a decreasing function in this region and initially $g_{1}(x)$ is smaller but sn increasing function, until it is larger than $\pi(x)$ and reaches its maximum where it then becomes a decreasing function but always larger than $\pi(x)$. Thus setting $M_{3}=\frac{\pi(s)}{g(s)}=\frac{a}{\lambda s}exp(\lambda |s-b|)$ will satisfy the equation $\pi(x) \leq M_{3}g(x)$ for all x in the region.

Subbing in the values of a=2, b=3 and s=1.25 given in the question the value of M for $\lambda=1$ is $M=max(\epsilon,\frac{e^{4.75}}{3},1.6e^{1.75}) \approx 38.53$ and for $\lambda=0.5$ is $M=max(\epsilon,\frac{e^{2.375}}{1.5},3.2e^{0.875})\approx 7.7$. Thus as M is smaller for $\lambda=0.5$. Thus, out of these 2 values for lambda 0.5 is my chosen value as it leads to an algorithm with a higher efficiency.

## Part b:
To implement A-R sampling using the proposal density with $\lambda=0.5$ to obtain samples from $\pi(x)$ I use the following algorithm:

To sample from $g(x)$ I use the inverse transformation technique by drawing u from a uniform distribution then if $u<0.5$ set $x=b+\frac{log(2u)}{\lambda}$ else set $x=b-\frac{2-2u}{\lambda}$. Then after obtaining x a draw from $g(x)$ perform A-R sampling by drawing y from a uniform(0,1) distribution and if $s < x$ then checking if $y \leq \frac{\pi(x)}{Mg(x)}=\frac{as^{a}}{2x^{a+1}}\frac{2}{M\lambda exp(-\lambda|x-b|)}$ else if $s-b <x$ then checking if $y \leq \frac{\pi(x)}{Mg(x)}=\frac{1}{2b}\frac{2}{M\lambda exp(-\lambda|x-b|)}$ if either of these checks are true then x is a draw from the target distribution $\pi(x)$ and we save this value, if these checks are not true or $x<s-b$ then x is not a draw from $\pi(x)$ and we throw away x. I repeat this process until the desired number of samples from $\pi(x)$ has been attained.

## Part d:
```{r,include=F}
#c
q1.ar <- function(n,a,b,s,lambda,M){

    X <- NULL
    for (i in 1:n){
      repeat{
        #Drawing x from the proposal distibution 
        u <- runif(1)
        if (u<0.5){
          x <- b+log(2*u)/lambda
        }else {
          x <- b-log(2-2*u)/lambda
        }
        #Performing A-R sampling
        y <- runif(1)
        if (s < x){
          if(y <= (a*s^a)/(2*x^(a+1))/(M*lambda*exp(-lambda*abs(x-b))/2)){
            X[i]<-x
            break
          }
        } else if(s-b<x){
          if(y <= 1/(2*b)/(M*lambda*exp(-lambda*abs(x-b))/2)){
            X[i]<-x
            break
          }
        }
      }
    }
    return(X)
}

#d
set.seed(2022)
ar.sample <- q1.ar(5000,2,3,1.25,0.5,M=3.2*exp(0.875))
#Creating function to create actual pdf of target distribution
target.dist <- function(x,a,b,s){
  n<-length(x)
  y<-rep(0,n)
  for (i in 1:n){

  if (x[i]<s-b){
    y[i] <- 0
  }else if (x[i]>s){
    y[i] <- (a*s^a)/(2*x[i]^(a+1))
  } else {
    y[i] <- 1/(2*b)
  }
  }
  return(y)
}
```

```{r,echo=F,fig.align='center'}
#Creating curve of target distribution
hist(ar.sample,nclass=100,prob = T,
     xlab = "Values generated by AR sampling",
     main = "Simulated and actual pdf of target distribution");
curve(target.dist(x,2,3,1.25),from=-4,to=20,n=1000,add=T, lwd= 1, lty= 1,col="red");
legend("topright", lty = c(1), cex = 0.6, col = c("red"),
legend = c("Actual pdf"))
```

```{r,include=F}
#e
#Using the plain Monte Carlo estimator of I (I_{n})

#Function to calculate the kth root of kth moment using plain Monte Carlo estimator
k.root.k.moment <- function(sample,n,k){
  output <- (1/n*sum(sample^k))^(1/k)
  return(output)
}

#k=2
round(k.root.k.moment(ar.sample,5000,2),3)
#2nd root of 2nd moment is 2.204 (3 d.p.)

#k=3
round(k.root.k.moment(ar.sample,5000,3),3)
#3rd root of 3rd moment is 3.350 (3 d.p.)
```

### 2022-12-17 draft_report.Rmd
Adding in graph for Q1

@@
 For multiple possible values of $\lambda$ in $g(x)$ the value of $\lambda$ that results in a smaller constant M will have the higher acceptance probability.
-To find M for each $\lambda$ value I split the real domain of x into 3 intervals corresponding to the different PARTS of $\pi$ and study the relationship between $g(x)$ and $\pi(x)$ within these 3 intervals to find $M_{1},M_{2},M_{3}$ which correspond to the smallest constant such that $\pi(x) \leq M_{i}g(x)$ for the ith interval.
+To find M for each $\lambda$ value I split the real domain of x into 3 regions (see plot 1) corresponding to the different PARTS of $\pi$ and study the relationship between $g(x)$ and $\pi(x)$ within these 3 regions to find $M_{1},M_{2},M_{3}$ which correspond to the smallest constant such that $\pi(x) \leq M_{i}g(x)$ for the ith interval.
 Then $M=max(M_{1},M_{2},M_{3})$ as this is the smallest constant that satisfies $\pi(x) \leq Mg(x)$ for all possible values of x.
@@
 For the second region [$s-b,s$], $\pi(x)=\frac{1}{2b}$.
-Considering that $g(x)$ is an increasing function in this region it will be at it's minimum value at $x=s-b$ and that the value of $\pi(x)$ is constant, setting $M_{2}=\frac{\pi(s-b)}{g(s-b)}=\frac{exp(\lambda|s-2b|)}{\lambda b}$ will satisfy the equation $\pi(x) \leq M_{2}g(x)$ for all x in the region.
+Considering that $g(x)$ is an increasing function in this region (see plot 1) it will be at it's minimum value at $x=s-b$ and that the value of $\pi(x)$ is constant, setting $M_{2}=\frac{\pi(s-b)}{g(s-b)}=\frac{exp(\lambda|s-2b|)}{\lambda b}$ will satisfy the equation $\pi(x) \leq M_{2}g(x)$ for all x in the region.
 For the third region ($s,\infty$), $\pi(x)=\frac{as^{a}}{2x^{a+1}}$.
-I note that $\pi(x)$ is a decreasing function in this region and initially $g_{1}(x)$ is smaller but sn increasing function, until it is larger than $\pi(x)$ and reaches its maximum where it then becomes a decreasing function but always larger than $\pi(x)$.
-Thus setting $M_{3}=\frac{\pi(s)}{g(s)}=\frac{a}{\lambda s}exp(\lambda |s-b|)$ will satisfy the equation $\pi(x) \leq M_{3}g(x)$ for all x in the region.
+I note that $\pi(x)$ is a decreasing function in this region (see plot 1) and initially $g(x)$ is a smaller but increasing function (see plot 1), until it reaches its maximum (larger than $\pi(x)$ at this value of x) where it then decreases for all larger x but always is larger than $\pi(x)$.
+Thus setting $M_{3}=\frac{\pi(s)}{g(s)}=\frac{a}{\lambda s}exp(\lambda |s-b|)$ will satisfy the equation $\pi(x) \leq M_{3}g(x)$ for all x in this region.
 
@@
 Thus, out of these 2 values for lambda 0.5 is my chosen value as it leads to an algorithm with a higher efficiency.
+
+```{r,echo=F,fig.align='center'}
+#a
+#Creating plots to help explain
+
+gx.1 <- function(x){
+  gx <- 0.5*exp(-abs(x-3))
+  return(gx)
+}
+
+gx.2 <- function(x){
+  gx <- 0.25*exp(-0.5*abs(x-3))
+  return(gx)
+}
+
+pi.x <- function(x){
+  pi <-NULL
+  for (i in 1:length(x)){
+  if (x[i]>1.25){
+    pi[i] <- (2*1.25^(2))/(2*x[i]^(3))
+  } else if (x[i] >= -1.75){
+    pi[i] <- 1/6
+  }else{
+    pi[i] <- 0
+  }
+  }
+  return(pi)
+}
+
+x<- seq(from=-4, to=10, by=0.01)
+
+plot(x=x,y=pi.x(x),
+     type="l",
+     ylab = "density",
+     main="Plot 1: Comparing target and proposal pdf's",
+     sub="Note: used a=2,b=3,s=1.25 with the green line seperating the regions of the graph")
+curve(gx.1(x),add=T,col="red")
+curve(gx.2,add=T,col="blue")
+abline(v=-1.75,lty=2,col="green")
+abline(v=1.25,lty=2,col="green")
+text(x=-3,y=0.4,labels="Region 1")
+text(x=0,y=0.4,labels="Region 2")
+text(x=6,y=0.4,labels="Region 3")
+legend("topright", lty = c(1,1,1), cex = 0.6, col = c("black","red","blue"),
+legend = c("Target distribution pdf","g(x) pdf with lambda=1","g(x) pdf with lambda=0.5"))
+```
 

### 2022-12-17 draft_report.Rmd
Starting on creating final R code script (Finished initial add of Q1,4 and 5)

@@
 
-pi.x <- function(x){
+target.dist<- function(x){
   pi <-NULL
@@
 
-plot(x=x,y=pi.x(x),
+plot(x=x,y=target.dist(x),
      type="l",

### 2022-12-17 Final.code.R

@@ -1,0 +1,345 @@
+#Setting up enviroment
+setwd("~/Desktop/Statistical Computing/CW/SC-CW2")
+
+######################### Question 1 ###############################
+
+### Part a ###
+
+#Creating plot to help explain choosing value of lambda with highest acceptance rate
+
+#Note assume a=2, b=3 and s=1.25 for all functions creating pdfs below
+
+#Function of pdf of proposal density with lambda=1
+gx.1 <- function(x){
+  gx <- 0.5*exp(-abs(x-3))
+  return(gx)
+}
+
+#Function of pdf of proposal density with lambda=0.5
+gx.2 <- function(x){
+  gx <- 0.25*exp(-0.5*abs(x-3))
+  return(gx)
+}
+
+#Creating function to create actual pdf of target distribution 
+target.dist<- function(x){
+  pi <-NULL
+  for (i in 1:length(x)){
+    if (x[i]>1.25){
+      pi[i] <- (2*1.25^(2))/(2*x[i]^(3))
+    } else if (x[i] >= -1.75){
+      pi[i] <- 1/6
+    }else{
+      pi[i] <- 0
+    }
+  }
+  return(pi)
+}
+
+#Creating plot
+x<- seq(from=-4, to=10, by=0.01)
+
+plot(x=x,y=target.dist(x),
+     type="l",
+     ylab = "density",
+     main="Plot 1: Comparing target and proposal pdf's",
+     sub="Note: used a=2,b=3,s=1.25 with the green line seperating the regions of the graph")
+curve(gx.1(x),add=T,col="red")
+curve(gx.2,add=T,col="blue")
+abline(v=-1.75,lty=2,col="green")
+abline(v=1.25,lty=2,col="green")
+text(x=-3,y=0.4,labels="Region 1")
+text(x=0,y=0.4,labels="Region 2")
+text(x=6,y=0.4,labels="Region 3")
+legend("topright", lty = c(1,1,1), cex = 0.6, col = c("black","red","blue"),
+       legend = c("Target distribution pdf","g(x) pdf with lambda=1","g(x) pdf with lambda=0.5"))
+
+### Part c ###
+
+q1.ar <- function(n,a,b,s,lambda,M){
+  #Creating vector to store values from target distribution
+  X <- NULL
+  for (i in 1:n){
+    repeat{
+      #Drawing samples x from the propsal distibution using inverse transformation technique
+      u <- runif(1)
+      if (u<0.5){
[…]

### 2022-12-20 Figuring_Q3.Rmd
Working on figuring out Q3 (using a equal to current proposal value)

@@
 ---
+```{r}
+# Part b
+calcDirichletPDF <- function(x,a){
+
+  #Calculating density of x under Dirichlet distribution with parameters specified by a
+  density <- (gamma(sum(a))/prod(gamma(a)))*prod(x^(a-1))
+  #calculating log of density
+  log.density <- log(density)
+  #returning a numeric vector
+  return(c(log.density))
+}
+
+# x <- PT[i,,j-1]
+# a <- 20*PT[i,,j-1]
+
+#Part c
+simDirichletRV <- function(a){
+  #Drawing from gamma distribution
+  U <- rgamma(a,1)
+  #Computing V a vector drawn from Dirichlet distribution with parameters specified by a
+  V <- U/sum(U)
+
+  return(V)
+}
+
+#Initial values
+X <- read.table("q3Weather.txt",header = F)
+X <- X$V1
+initPT <- matrix(1/3,nrow=3,ncol=3)
+initStateProbs <- c(1/3,1/3,1/3)
+n <- 100000
+burnin <- 10000
+
+#Converting X into a numeric vector to access entries of PT
+X <- ifelse(X=="s",1,(ifelse(X=="c",2,3)))
+init.state <- X[1:(length(X)-1)]
+trans.state <- X[2:length(X)]
+
+#initializing algorithm
+PT <- array(NA,dim=c(3,3,n+burnin+1))
+PT[,,1] <- initStateProbs
+proposal <- matrix(NA,ncol=3,nrow=3)
+
+#Starting loop 
+for(j in 2:(n+burnin+1)){
+
+#Generating proposal value for each row of transition matrix (each weather state p_{s},p_{c},p_{r})
+for (i in 1:length(initStateProbs)){
+  proposal[i,] <- simDirichletRV(20*PT[i,,j-1])
+}
+
+#Generating posterior using equation (12) for proposal (new) and current value of PT
+posterior.cur <- posterior.new <- NULL
+for (i in 1:length(init.state)){
+  posterior.cur[i] <- PT[init.state[i],trans.state[i],j-1]
+  posterior.new[i] <- proposal[init.state[i],trans.state[i]]
+}
+# Calculating density of posterior using equation 12
+posterior.density.cur <- prod(posterior.cur)*initStateProbs[X[1]]
+posterior.density.new <-prod(posterior.new)*initStateProbs[X[1]]
+
+# Calculating density using proposal density for each row (p_{x})
+proposal.density.cur <- proposal.density.new <- NULL
+for (i in 1:length(initStateProbs)){
+proposal.density.new[i] <- calcDirichletPDF(proposal[i,],20*PT[i,,j-1])
+proposal.density.cur[i] <- calcDirichletPDF(PT[i,,j-1],20*PT[i,,j-1])
+}
+#How do we combine the 3 vectors to compare with PT??? - do we just multiple to get joint probability density
+proposal.density.cur <- exp(sum(proposal.density.cur))
+proposal.density.new <- exp(sum(proposal.density.new))
+
+#accept-reject
+alpha <- (posterior.density.new*proposal.density.cur)/(posterior.density.cur*proposal.density.new)
+
+if (runif(1) <= min(1,alpha) ){
+  PT[,,j] <- proposal
+}else{
+  PT[,,j] <- PT[,,j-1]
+}
+
+}
+
+######### Testing matrox selection
+# PT[,,1] <- initPT <- matrix(c(1/2,0.25,0.25,1/3,1/6,0.5,1/8,1/4,5/8),nrow=3,ncol=3,byrow = T)
+# PT[2,,1]
+
+########
+```
+
+```{r}
+
+# Part b
+calcDirichletPDF <- function(x,a){
+
+  #Calculating density of x under Dirichlet distribution with parameters specified by a
+  density <- (gamma(sum(a))/prod(gamma(a)))*prod(x^(a-1))
+  #calculating log of density
+  log.density <- log(density)
+  #returning a numeric vector
+  return(c(log.density))
+}
+
+#Part c
+simDirichletRV <- function(a){
+  #Drawing from gamma distribution
+  U <- rgamma(a,1)
+  #Computing V a vector drawn from Dirichlet distribution with parameters specified by a
+  V <- U/sum(U)
+
+  return(V)
+}
+
+#Initial values
+X <- read.table("q3Weather.txt",header = F)
+X <- X$V1
+initPT <- matrix(1/3,nrow=3,ncol=3)
+initStateProbs <- c(1/3,1/3,1/3)
+n <- 100000
+burnin <- 10000
+
+q3.mcmc <- function(X,initStateProbs,initPT,burnin,n){
+#Converting X into a numeric vector to access entries of PT
+X <- ifelse(X=="s",1,(ifelse(X=="c",2,3)))
+init.state <- X[1:(length(X)-1)]
+trans.state <- X[2:length(X)]
+
+#initializing algorithm
+PT <- array(NA,dim=c(3,3,n+burnin+1))
+PT[,,1] <- initStateProbs
+proposal <- matrix(NA,ncol=3,nrow=3)
+
+#Starting loop 
+for(j in 2:(n+burnin+1)){
+
+  #Generating proposal value for each row of transition matrix (each weather state p_{s},p_{c},p_{r})
+  for (i in 1:length(initStateProbs)){
+    proposal[i,] <- simDirichletRV(20*PT[i,,j-1])
+  }
+
+  #Generating posterior using equation (12) for proposal (new) and current value of PT
+  posterior.cur <- posterior.new <- NULL
+  for (i in 1:length(init.state)){
+    posterior.cur[i] <- PT[init.state[i],trans.state[i],j-1]
+    posterior.new[i] <- proposal[init.state[i],trans.state[i]]
+  }
+  # Calculating density of posterior using equation 12
+  posterior.density.cur <- prod(posterior.cur)*initStateProbs[X[1]]
+  posterior.density.new <-prod(posterior.new)*initStateProbs[X[1]]
+
+  # Calculating density using proposal density for each row (p_{x})
+  proposal.density.cur <- proposal.density.new <- NULL
+  for (i in 1:length(initStateProbs)){
+    proposal.density.new[i] <- calcDirichletPDF(proposal[i,],20*PT[i,,j-1])
+    proposal.density.cur[i] <- calcDirichletPDF(PT[i,,j-1],20*PT[i,,j-1])
+  }
+  #How do we combine the 3 vectors to compare with PT??? - do we just multiple to get joint probability density
+  proposal.density.cur <- exp(sum(proposal.density.cur))
+  proposal.density.new <- exp(sum(proposal.density.new))
+
+  #accept-reject
+  alpha <- (posterior.density.new*proposal.density.cur)/(posterior.density.cur*proposal.density.new)
+
+  if (runif(1) <= min(1,alpha) ){
+    PT[,,j] <- proposal
+  }else{
+    PT[,,j] <- PT[,,j-1]
+  }
+
+}
+#Removing burnin samples
+return(PT[,,(burnin+1):(n+burnin+1)])
+}
+
+set.seed(2022)
+
+PT.sample <- q3.mcmc(X,initStateProbs,initPT,burnin,n)
+
+table(X)
+
+PT.expected <- matrix(c(mean(PT.sample[1,1,]),mean(PT.sample[1,2,]),mean(PT.sample[1,3,]),
+                        mean(PT.sample[2,1,]),mean(PT.sample[2,2,]),mean(PT.sample[2,3,]),
+                        mean(PT.sample[3,1,]),mean(PT.sample[3,2,]),mean(PT.sample[3,3,])),ncol=3,nrow=3,byrow = T)
+round(PT.expected,3)
+```
+
 ```{r}
@@
 ```
+
+```{r}
+## Trying again monday 19th dec
+
+###Testing 
+#Initial values
+#finding prob of sequence under current pt
+X <- read.table("q3Weather.txt",header = F)
+X <- X$V1
+initPT <- matrix(1/3,nrow=3,ncol=3)
+initStateProbs <- c(1/3,1/3,1/3)
+n <- 100000
+burnin <- 500
+
+  #Converting X into a numeric vector to access entries of PT
+  X <- ifelse(X=="s",1,(ifelse(X=="c",2,3)))
+  init.state <- X[1:(length(X)-1)]
+  trans.state <- X[2:length(X)]
+
+  #initializing algorithm
+  V <- array(NA,dim=c(3,3,n+burnin+1))
+  V[,,1] <- initStateProbs
+  est.state.prob <- array(NA,dim=c(1,3,n+burnin+1)) 
+  est.state.prob[,,1] <-  initStateProbs
+
+proposal <- matrix(NA,ncol=3,nrow=3)
+#starting loop
+for (j in 2:(n+burnin+1)){
+
+    est.state.prob[,,j] <- est.state.prob[,,j-1]%*%V[,,j-1]
+    #Generating proposal value for each row of transition matrix
+    for (i in 1:length(initStateProbs)){
+      proposal[i,] <- simDirichletRV(20*V[i,,j-1])
+    }
+     #calculating alpha
+
+      #Generating posterior using equation (12) for proposal and old value of PT
+      actualp.prev <- actualp.prop <- NULL
+      for (i in 1:length(init.state)){
+        actualp.prev[i] <- V[init.state[i],trans.state[i],j-1]
+        actualp.prop[i] <- proposal[init.state[i],trans.state[i]]
+      }
+      #density has some constant as proportional but due to fraction constant cancels out so dont
+      #need to worry about it
+      true.density.prev <- prod(actualp.prev)*initStateProbs[X[1]]
+      true.density.prop <-prod(actualp.prop)*initStateProbs[X[1]]
+
+      #Finding density of propoisal using Diriclet(20*prev.value)
+      prop.dens.prop <- prop.dens.prev <- NULL
+       for (i in 1:length(initStateProbs)){
+        prop.dens.prop[i] <- calcDirichletPDF(proposal[i,],20*V[i,,j-1])
+        prop.dens.prev[i] <- calcDirichletPDF(20*V[i,,j-1],20*V[i,,j-1])
+
+       }
+      total.prop.dens.prop <- exp(sum(prop.dens.prop))
+      total.prop.dens.prev <- exp(sum(prop.dens.prev))
+
+      # alpha <- (true.density.prop*prop.density.prev)/(true.density.prev*prop.density.prop)
+      alpha <-  (true.density.prop/true.density.prev)*(total.prop.dens.prev/total.prop.dens.prop)
+
+      if(min(1,alpha)>=runif(1)){
+        V[,,j] <- proposal
+      }else{
+        V[,,j] <- V[,,j-1]
+      }
+}
+```

### 2022-12-21 draft_report.Rmd
Adding in draft of explanation of Q2

@@
 ```
+
+# Question 2
+
+## Part a
+For the continuous-time random walk conditioned on $U_{0}=u_{0}$
+\begin{equation}
+U_{i} \sim \text{Normal}(\mu=U_{i-1},\sigma^{2}=t_{i}-t_{i-1})
+\end{equation}
+
+First by transforming $a$ to the logit scale by setting $b_{i}=log(\frac{a_{i}}{1-a_{i}})$ $p_{tail}(a)=Pr(X>a|X_{0}=x_{0})  \equiv Pr(U>b|U_{0}=u_{0}=p_{tail}(b))$, thus we can work with the sequence $U_{i}$ to answer equation (6).
+Now from equation (3):
+\begin{equation*}
+Pr(U_{i}>b_{i})=Pr((\sqrt{t_{i}-t_{i-1}})Z+u_{i-1}>b_{i}))=Pr\left(Z>\frac{b_{i}-u_{i-1}}{\sqrt{t_{i}-t_{i-1}}}\right) \hspace{2em}  
+\end{equation*}
+For Z the standard normal and $u_{i-1}$ a sample from $U_{i-1}$
+Therefore, to not waste any samples let $q_{i}$ the proposal the ith observation in the sequence $U_{i}$ be an exponential distribution shifted by $s_{i}$ where:
+
+\begin{equation*}
+ s_{i}=\frac{b_{i}-u_{i-1}}{\sqrt{t_{i}-t_{i-1}}} 
+\end{equation*}
+
+As the samples $w=\{w_{1},...,w_{d}\}$ drawn from these proposals $q=\{q_{1},....,q_{d}\}$ will satisfy $w_{i}>b_{i},\hspace{1em} i \in \{1,...,d\}$ thus all values can be used to calculate $p_{tail}(b)=p_{tail}(a)$.
+
+To calculate $q_{i}$ however, we need to know the value $u_{i-1}$ (a value drawn from the random variable $U_{i-1}$).
+Thus, each iteration will need to generate these $u_{i}$ values using equation (3) the initial value $u_{0}$ and times $t_{i}i \in \{0,...,d\}$ to generate a realization $u$ of the random walk sequence $U$ and storing these in a vector to access to generate a sample from the proposal distribution.
+
+## Part b
+
+The algorithm will start by applying the logit transform to a and then using these transformed values b to work with values in sequence
+
+First by transforming $\boldsymbol{a}$ to the logit scale by setting $b_{i}=log(\frac{a_{i}}{1-a_{i}})$ $p_{tail}(\boldsymbol{a})=Pr(\boldsymbol{X}>\boldsymbol{a}|X_{0}=x_{0})  \equiv Pr(\boldsymbol{U}>\boldsymbol{b}|U_{0}=u_{0})=p_{tail}(\boldsymbol{b}))$, thus we can work with the sequence $U_{i}$ to answer equation (6).
+Therefore start the algorithm by transforming inputs into proportions (if given as percentages) and calculating $\boldsymbol{b}$.
+
+Next we must simulate the random walk $U_{i}$ using equation 3 and storing these realizations $u_{i}$ in a vector.
+Using these values and the proposals $q_{i}$ in part a generate $v_{i}$ a value for each step in sequence $U_{i}$
+Calculate weights $w_{i}$ for each $v_{i}$ by formula $w_{i}=\frac{\pi_{i}(v_{i})}{Q_{i}(v_{i})})$ where $\pi_{i}$ is the pdf of $U_{i}$ and $Q_{i}$ is the density of the proposal distribution $q_{i}$
+Then to calculate equation (6) for this iteration find the product of the weights as (put in equation with integral and how sampled pop is that integral)
+Repeat this process until have desired n values and then calculate importance sampling estimator by finding mean of these values.

### 2022-12-21 draft1_Rcode.Rmd
Updating final R code to include Q2 code and part b and c from Q3

@@
 ## Question 2
-n <- 100000
-X <- rexp(n,rate=0.5)+30 ## generate samples from shifted Exp(10/7)
-v.weights <- dnorm(X,mean=30,sd=0.7)/dexp(X-30,rate=0.5) ## vector of weights
-p.t <- mean(v.weights) ## calculate the IS estimator
 
-q2.is <- function(n,x0,time,a,lambda){
+# Trial function calculates probability of X_1 >a_1 given initial value (need to now work out how to do the rest)
+q2.is.trial <- function(n,x0,time,a,lambda){
+  #Turning values into proportions
+  a <- a/100
+  x0 <- x0/100
+  #transforming a to logit scale and turning a into a proportion (dividing by 100)
+  b <- log((a)/(1-a))
+  #generating samples from shifted exponential
+  X <- rexp(n,rate = lambda) + (b-x0)/sqrt(time[2]-time[1])
+  #Calculating weights
+  v.weights <- dnorm(X)/dexp(X-((b-x0)/sqrt(time[2]-time[1])),rate=lambda)
+  #Finding I.S estimator
+  p.t <- mean(v.weights)
+  return(p.t)
+}
 
+q2.is.trial(n=10000,x0=30,time=c(0,0.7),a=50,lambda=0.5)
+
+q2.is.trial.chain <- function(n,x0,time,a,lambda){
+  #Turning values into proportions
+  a <- a/100
+  x0 <- x0/100
+  #transforming a to logit scale (calling this vector b)
+  b <- log((a)/(1-a))
+
+  p.b.iter <- NULL
+
+  if (length(time)>2){  #Case when Markov chain has more than 1 transition thus have to simulate a random walk to calculate the proposal distribution
+
+    for (k in 1:n){
+
+      #Simulating random walk
+      r.w <- NULL
+      r.w[1] <- x0
+      for (i in 2:(length(time)-1)){
+        r.w[i] <- rnorm(1,mean=r.w[i-1],sd=sqrt(time[i]-time[i-1]))
       }
+
+      #generating samples from shifted exponential
+      X <- NULL
+      for (i in 2:length(time)){
+        X[i-1] <- rexp(1,rate = lambda) + (b[i-1]-r.w[i-1])/sqrt(time[i]-time[i-1])
+      }
+      #Calculating weights for each observation in random walk
+      v.weights <- NULL
+      for (i in 2:length(time)){
+        v.weights[i-1] <- dnorm(X[i-1])/dexp(X[i-1]-((b[i-1]-r.w[i-1])/sqrt(time[i]-time[i-1])),rate=lambda)
+      }
+      #Finding probability that X is greater than a (satisfies equation 6) for current iteration 
+      p.b.iter[k] <- prod(v.weights)
+      }
+      #Finding I.S estimator by taking mean 
+      I.S.est <- mean(p.b.iter)
+
+  } else { # Dealing with case that markov chain only has 1 jump so don't need to calculate random walk
+           # Simplifies to case that X_{1}>a given that X_{0}=x0
+
+    #generating samples from proposal shifted exponential
+    X <- rexp(n,rate = lambda) + (b-x0)/sqrt(time[2]-time[1])
+    #Calculating weights
+    v.weights <- dnorm(X)/dexp(X-((b-x0)/sqrt(time[2]-time[1])),rate=lambda)
+    #Finding I.S estimator
+    I.S.est  <- mean(v.weights)
+  }
+
+  return(I.S.est)
+}
+set.seed(2022)
+q2.is.trial.chain(n=10000,x0=30,time=c(0,0.7,1.2,2.5),a=c(50,70,90),lambda=0.5)
+set.seed(2022)
+q2.is.trial.chain(n=10000,x0=30,time=c(0,0.7),a=c(50),lambda=0.5)
+
+set.seed(2022)
+n=10000
+x0=5
+time=c(0,0.7,1.2,2.5)
+a=c(50,70,90)
+lambda=0.5
+
+  a <- a/100
+  x0 <- x0/100
+  #transforming a to logit scale and turning a into a proportion (dividing by 100)
+  b <- log((a)/(1-a))
+  p.t <- NULL
+  for (k in 1:n){
+
+  #Simulating random walk
+  r.w <- NULL
+  r.w[1] <- x0
+  for (i in 2:(length(time)-1)){
+    r.w[i] <- rnorm(1,mean=r.w[i-1],sd=sqrt(time[i]-time[i-1]))
+  }
+
+  #generating samples from shifted exponential
+  X <- NULL
+  for (i in 2:length(time)){
+  X[i-1] <- rexp(1,rate = lambda) + (b[i-1]-r.w[i-1])/sqrt(time[i]-time[i-1])
+  }
+  #Calculating weights for each observation in random walk
+  v.weights <- NULL
+  for (i in 2:length(time)){
+  v.weights[i-1] <- dnorm(X[i-1])/dexp(X[i-1]-((b[i-1]-r.w[i-1])/sqrt(time[i]-time[i-1])),rate=lambda)
+  }
+  #Finding I.S estimator (product means to find overall)
+  p.t[k] <- prod(mean(v.weights))
+  }
+
+n <- 100
+x0 <- 30
+time <- c(0,0.7)
+a <- 50
+lambda <- 0.5
+
+  b <- log((a/100)/(1-(a/100)))
+  #generating samples from shifted exponential
+  X <- rexp(n,rate = lambda) + (b-x0)/sqrt(time[2]-time[1])
+  v.weights <- dnorm(X)/dexp(X-((b-x0)/sqrt(time[2]-time[1])),rate=lambda)
+  p.t <- mean(v.weights)
+  return(p.t)
 ```

### 2022-12-21 Final.code.R

@@ -123,6 +123,93 @@ round(k.root.k.moment(ar.sample,5000,3),3)
 #3rd root of 3rd moment is 3.350 (3 d.p.)
 
 
+########################################## Question 2 #################################################
+
+### Part c ###
+
+q2.is <- function(n,x0,time,a,lambda){
+  #Turning values into proportions
+  a <- a/100
+  x0 <- x0/100
+  #transforming a to logit scale (calling this vector b)
+  b <- log((a)/(1-a))
+  
+  p.b.iter <- NULL
+  
+  if (length(time)>2){  #Case when Markov chain has more than 1 transition (d>1) thus have to simulate a random walk to calculate the proposal distribution 
+    
+    for (k in 1:n){
+      
+      #Simulating random walk
+      r.w <- NULL
+      r.w[1] <- x0
+      for (i in 2:(length(time)-1)){
+        r.w[i] <- rnorm(1,mean=r.w[i-1],sd=sqrt(time[i]-time[i-1]))
+      }
+      
+      #generating samples from shifted exponential
+      X <- NULL
+      for (i in 2:length(time)){
+        X[i-1] <- rexp(1,rate = lambda) + (b[i-1]-r.w[i-1])/sqrt(time[i]-time[i-1])
+      }
+      #Calculating weights for each observation in random walk
+      v.weights <- NULL
+      for (i in 2:length(time)){
+        v.weights[i-1] <- dnorm(X[i-1])/dexp(X[i-1]-((b[i-1]-r.w[i-1])/sqrt(time[i]-time[i-1])),rate=lambda)
+      }
+      #Finding probability that X is greater than a (satisfies equation 6) for current iteration 
+      p.b.iter[k] <- prod(v.weights)
+    }
+    #Finding IS estimator by taking mean 
+    I.S.est <- mean(p.b.iter)
+    
+  } else { # Dealing with case that markov chain only has 1 jump so don't need to calculate random walk
+    # Simplifies to case that X_{1}>a given that X_{0}=x0
+    
+    #generating samples from proposal shifted exponential
+    X <- rexp(n,rate = lambda) + (b-x0)/sqrt(time[2]-time[1])
+    #Calculating weights
+    v.weights <- dnorm(X)/dexp(X-((b-x0)/sqrt(time[2]-time[1])),rate=lambda)
+    #Finding I.S estimator
+    I.S.est  <- mean(v.weights)
+  }
+  
+  return(I.S.est)
+}
+
+
[…]

### 2022-12-21 draft_report.Rmd
Adding in Q3 part a write up

@@
 
-Next we must simulate the random walk $U_{i}$ using equation 3 and storing these realizations $u_{i}$ in a vector.
+Next for each iteration j we must simulate the random walk $U_{i}$ using equation 3 and storing these realizations $u_{i}$ in a vector.
 Using these values and the proposals $q_{i}$ in part a generate $v_{i}$ a value for each step in sequence $U_{i}$
 Calculate weights $w_{i}$ for each $v_{i}$ by formula $w_{i}=\frac{\pi_{i}(v_{i})}{Q_{i}(v_{i})})$ where $\pi_{i}$ is the pdf of $U_{i}$ and $Q_{i}$ is the density of the proposal distribution $q_{i}$
-Then to calculate equation (6) for this iteration find the product of the weights as (put in equation with integral and how sampled pop is that integral)
-Repeat this process until have desired n values and then calculate importance sampling estimator by finding mean of these values.
+Then to calculate equation (6) for this iteration (and thus this particular simulated realization of the random walk) find the product of the weights as (put in equation with integral and how sampled pop is that integral)
+Repeat this iterative process until have desired n values and then calculate importance sampling estimator by finding mean of these values.
+(SAY SOMETHING ABOUT FACT THAT LARGE SAMPLE MEANS VALUE CONVERGES TO P_TAIL [AS TAKE OVER LOTS OF REALISATION OF RW])
+
+# Question 3
+
+## Part a
+At initialization set $P_{T}^{cur}$ to be initial value of $P_{T}$ and start at iteration j=1
+
+For each iteration j start by drawing $P_{T}^{new}$ the proposed value of $P_{T}$ from the proposal density by drawing each $p_{x}'$ with $x \in \{s,c,r\}$ from Dirichlet $(20p_{x}^{(j-1)})$ (use function created in part c for this) and combining to get $P_{T}^{new}=(p_{s}',p_{c}',p_{r}')^{T}$.
+
+Then calculate the ratio of the posterior density using the new and current values of $P_{T}$. First by considering that:
+
+\begin{equation}
+ p(P_{T}|X,\pi) \propto p(X|P_{T},\pi)
+\end{equation}
+
+We get $\frac{p(P_{T}^{new}|X,\pi)}{p(P_{T}^{(j-1)}|X,\pi)}=\frac{p(X|P_{T}^{new},\pi)}{p(X|P_{T}^{j-1},\pi)}$, thus we can calculate this ratio of densities as we can calculate the RHS.
+
+To calculate the RHS using the Markov property and equation 8 and 9 in the coursework question sheet:
+
+\begin{equation}
+Pr(X=x|\pi,P_{T}) = Pr(X_{1}=x_{1}|\pi,,P_{T})Pr(X_{2}=x_{2}|X_{1}=x_{1},,P_{T})...Pr(X_{n}=x_{n}|X_{n-1}=x_{n-1},,P_{T}) = \pi_{x_{1}}\prod_{i=2}^{n}p_{T}(X_{i}=x_{i}|X_{i-1}=x_{i-1}
+\end{equation}
+
+Now to calculate the ratio of the proposal densities (with the proposal density denoted with $q$) using current and new values of $P_{T}$.
+To do this, find the densities of each row of $P_{T}$ using the function created in part b with the input x the values of $p_{x}^{new}$ or $p_{x}^{j-1}$ and a=$20p_{x}^{j-1}$ then to calculate the joint probability by multiplying the densities for each row together to find $q(P_{T}^{new})$ and $q(P_{T}^{j-1})$
+
+Then calculate $\alpha = \frac{p(P_{T}^{new}|X,\pi)}{p(P_{T}^{j-1}|X,\pi)}\frac{q(P_{T}^{j-1})}{q(P_{T}^{new})}$ and generate $u$ a value drawn from a Uniform(0,1) distribution and if $u<=min(1,\alpha)$ set $P_{T}^{new}$ to be the jth value of $P_{T}$.
+Otherwise, set $P_{T}^{cur}$ to be jth value of $P_{T}$.
+
+Once generated the desired number of $P_{T}$ throw away the first b values where b is the burn in. The remaining values are the estimated posterior distribution of $P_{T}$.

### 2022-12-21 Figuring_Q3.Rmd
Adding in code for part d and e of Q3 (MCMC question)

@@
 }
-#How do we combine the 3 vectors to compare with PT??? - do we just multiple to get joint probability density
+#Finding probability of P_T using joint probability 
 proposal.density.cur <- exp(sum(proposal.density.cur))

### 2022-12-21 Final.code.R

@@ -210,6 +210,86 @@ simDirichletRV <- function(a){
   return(V)
 }
 
+
+### Part d ###
+
+q3.mcmc <- function(X,initStateProbs,initPT,burnin,n){
+  #Converting X into a numeric vector to access entries of PT
+  X <- ifelse(X=="s",1,(ifelse(X=="c",2,3)))
+  init.state <- X[1:(length(X)-1)]
+  trans.state <- X[2:length(X)]
+  
+  #initializing algorithm
+  PT <- array(NA,dim=c(3,3,n+burnin+1))
+  PT[,,1] <- initStateProbs
+  proposal <- matrix(NA,ncol=3,nrow=3)
+  
+  #Starting loop 
+  for(j in 2:(n+burnin+1)){
+    
+    #Generating proposal value for each row of transition matrix (each weather state p_{s},p_{c},p_{r})
+    for (i in 1:length(initStateProbs)){
+      proposal[i,] <- simDirichletRV(20*PT[i,,j-1])
+    }
+    
+    #Generating posterior using equation (12) for proposal (new) and current value of PT
+    posterior.cur <- posterior.new <- NULL
+    for (i in 1:length(init.state)){
+      posterior.cur[i] <- PT[init.state[i],trans.state[i],j-1]
+      posterior.new[i] <- proposal[init.state[i],trans.state[i]]
+    }
+    # Calculating density of posterior using equation 12
+    posterior.density.cur <- prod(posterior.cur)*initStateProbs[X[1]]
+    posterior.density.new <-prod(posterior.new)*initStateProbs[X[1]]
+    
+    # Calculating density using proposal density for each row (p_{x})
+    proposal.density.cur <- proposal.density.new <- NULL
+    for (i in 1:length(initStateProbs)){
+      proposal.density.new[i] <- calcDirichletPDF(proposal[i,],20*PT[i,,j-1])
+      proposal.density.cur[i] <- calcDirichletPDF(PT[i,,j-1],20*PT[i,,j-1])
+    }
+    #How do we combine the 3 vectors to compare with PT??? - do we just multiple to get joint probability density
+    proposal.density.cur <- exp(sum(proposal.density.cur))
+    proposal.density.new <- exp(sum(proposal.density.new))
+    
+    #accept-reject
[…]

### 2022-12-21 Figuring_Q3.Rmd
Changing Q3 part b

@@
 ---
+
+```{r}
+
+### Part b ###
+
+calcDirichletPDF <- function(x,a){
+
+  #Calculating density of x under Dirichlet distribution with parameters specified by a
+  density <- (gamma(sum(a))/prod(gamma(a)))*prod(x^(a-1))
+  #calculating log of density
+  log.density <- log(density)
+  #returning a numeric vector
+  return(c(log.density))
+}
+
+calcDirichletPDF(c(0.2,0.3,0.5),c(3,4,5))
+
+calcDirichletPDF.2 <- function(x,a){
+  #Calculating density of x under Dirichlet distribution with parameters specified by a
+  log.density <- lgamma(sum(a))-sum(lgamma(a))+sum((a-1)*log(x))
+  return(c(log.density))
+}
+
+calcDirichletPDF.2(c(0.2,0.3,0.5),c(3,4,5))
+
+### Part c ###
+
+simDirichletRV <- function(a){
+  #Drawing from gamma distribution
+  U <- rgamma(a,1)
+  #Computing V a vector drawn from Dirichlet distribution with parameters specified by a
+  V <- U/sum(U)
+
+  return(V)
+}
+
+### Part d ###
+
+q3.mcmc <- function(X,initStateProbs,initPT,burnin,n){
+  #Converting X into a numeric vector to access entries of PT
+  X <- ifelse(X=="s",1,(ifelse(X=="c",2,3)))
+  init.state <- X[1:(length(X)-1)]
+  trans.state <- X[2:length(X)]
+
+  #initializing algorithm
+  PT <- array(NA,dim=c(3,3,n+burnin+1))
+  PT[,,1] <- initStateProbs
+  proposal <- matrix(NA,ncol=3,nrow=3)
+
+  #Starting loop 
+  for(j in 2:(n+burnin+1)){
+
+    #Generating proposal value for each row of transition matrix (each weather state p_{s},p_{c},p_{r})
+    for (i in 1:length(initStateProbs)){
+      proposal[i,] <- simDirichletRV(20*PT[i,,j-1])
+    }
+
+    #Generating posterior using equation (12) for proposal (new) and current value of PT
+    posterior.cur <- posterior.new <- NULL
+    for (i in 1:length(init.state)){
+      posterior.cur[i] <- PT[init.state[i],trans.state[i],j-1]
+      posterior.new[i] <- proposal[init.state[i],trans.state[i]]
+    }
+    # Calculating density of posterior using equation 12
+    posterior.density.cur <- prod(posterior.cur)*initStateProbs[X[1]]
+    posterior.density.new <-prod(posterior.new)*initStateProbs[X[1]]
+
+    # Calculating density using proposal density for each row (p_{x})
+    proposal.density.cur <- proposal.density.new <- NULL
+    for (i in 1:length(initStateProbs)){
+      proposal.density.new[i] <- calcDirichletPDF(proposal[i,],20*PT[i,,j-1])
+      proposal.density.cur[i] <- calcDirichletPDF(PT[i,,j-1],20*PT[i,,j-1])
+    }
+    #How do we combine the 3 vectors to compare with PT??? - do we just multiple to get joint probability density
+    proposal.density.cur <- exp(sum(proposal.density.cur))
+    proposal.density.new <- exp(sum(proposal.density.new))
+
+    #accept-reject
+    alpha <- (posterior.density.new*proposal.density.cur)/(posterior.density.cur*proposal.density.new)
+
+    if (runif(1) <= min(1,alpha) ){
+      PT[,,j] <- proposal
+    }else{
+      PT[,,j] <- PT[,,j-1]
+    }
+
+  }
+  #Removing burn in samples
+  return(PT[,,(burnin+1):(n+burnin+1)])
+}
+
+#Initial values
+X <- read.table("q3Weather.txt",header = F)
+X <- X$V1
+initPT <- matrix(1/3,nrow=3,ncol=3)
+initStateProbs <- c(1/3,1/3,1/3)
+n <- 100000
+burnin <- 10000
+
+set.seed(2022)
+
+PT.sample <- q3.mcmc(X,initStateProbs,initPT,burnin,n)
+
+PT.sample <- q3.mcmc(rep("c",1000),initStateProbs,initPT,burnin,n)
+
+PT.expected <- matrix(c(mean(PT.sample[1,1,]),mean(PT.sample[1,2,]),mean(PT.sample[1,3,]),
+                        mean(PT.sample[2,1,]),mean(PT.sample[2,2,]),mean(PT.sample[2,3,]),
+                        mean(PT.sample[3,1,]),mean(PT.sample[3,2,]),mean(PT.sample[3,3,])),ncol=3,nrow=3,byrow = T)
+round(PT.expected,3)
+
+plot(PT.sample[2,2,],type="l")
+
+acf(PT.sample[2,2,])
+#autocorrelation goes to nothing very quickly :)
+
+hist(PT.sample,breaks=50)
+#Looks like distrbution so maybe isn't broken :)
+```
+
 ```{r}

### 2022-12-21 Final.code.R

@@ -190,12 +190,8 @@ round(is.estimate,4)
 ### Part b ###
 
 calcDirichletPDF <- function(x,a){
-  
-  #Calculating density of x under Dirichlet distribution with parameters specified by a
-  density <- (gamma(sum(a))/prod(gamma(a)))*prod(x^(a-1))
-  #calculating log of density
-  log.density <- log(density)
-  #returning a numeric vector
+  #Calculating log density of x under Dirichlet distribution with parameters specified by a
+  log.density <- lgamma(sum(a))-sum(lgamma(a))+sum((a-1)*log(x))
   return(c(log.density))
 }
 
@@ -253,9 +249,9 @@ q3.mcmc <- function(X,initStateProbs,initPT,burnin,n){
     proposal.density.new <- exp(sum(proposal.density.new))
     
     #accept-reject
-    alpha <- (posterior.density.new*proposal.density.cur)/(posterior.density.cur*proposal.density.new)
+    ratio <- (posterior.density.new*proposal.density.cur)/(posterior.density.cur*proposal.density.new)
     
-    if (runif(1) <= min(1,alpha) ){
+    if (runif(1) <= min(1,ratio) ){
       PT[,,j] <- proposal
     }else{
       PT[,,j] <- PT[,,j-1]

### 2022-12-23 draft_report.Rmd
Adding in code for part d and e of Q3 (MCMC question)
Updating report so it makes sense (or at least more sense :P)

--- 2022-12-17 draft_report.Rmd
@@
 ---
-title: "STatistical Computing (MATH6173) Coursework 2"
+title: "Statistical Computing (MATH6173) Coursework 2"
 author: "Student ID: 34273638"
@@
 ```{r, include=F}
-#Setting up enviroment
+#Setting up environment
 setwd("~/Desktop/Statistical Computing/CW/SC-CW2")
@@
 
-## Part a: (add graphs to help in explanation)
+## Part a:
 
@@
 
-For acceptance-rejection sampling we need to find constant $M>0$ such that $\pi(x) \leq Mg(x)$ for $\pi(x)$ the target distribution and $g(x)$ the proposal density.
-This constant M determines the acceptance probability and hence the efficiency of the algorithm.
-With the acceptance probability (denoted $\tau$) equal to $\frac{1}{M}$.
-For multiple possible values of $\lambda$ in $g(x)$ the value of $\lambda$ that results in a smaller constant M will have the higher acceptance probability.
-To find M for each $\lambda$ value I split the real domain of x into 3 regions (see plot 1) corresponding to the different PARTS of $\pi$ and study the relationship between $g(x)$ and $\pi(x)$ within these 3 regions to find $M_{1},M_{2},M_{3}$ which correspond to the smallest constant such that $\pi(x) \leq M_{i}g(x)$ for the ith interval.
-Then $M=max(M_{1},M_{2},M_{3})$ as this is the smallest constant that satisfies $\pi(x) \leq Mg(x)$ for all possible values of x.
+For acceptance-rejection sampling (AR sampling) we need to find a constant $M>0$ such that $\pi(x) \leq Mg(x)$ for $\pi(x)$ the target distribution and $g(x)$ the proposal density defined in equations 1 and 2 above.
+This constant M determines the acceptance probability which is $\frac{1}{M}$ and hence the efficiency of the algorithm.
+For multiple possible values of $\lambda$ in $g(x)$ the value of $\lambda$ that results in a smaller constant M will have a higher acceptance probability and thus be our preferred value of $\lambda$ for the algorithm.
+To find M for each $\lambda$ value I split the real domain of x into 3 regions (see plot 1) that I call 1,2 and 3 which correspond to the different expressions of $\pi$ in different regions and study the relationship between $g(x)$ and $\pi(x)$ within these 3 regions to find $M_{1},M_{2},M_{3}$ which correspond to the smallest constant such that $\pi(x) \leq M_{i}g(x)$ for the ith region.
+Then set $M=max(M_{1},M_{2},M_{3})$ as this is the smallest constant that satisfies $\pi(x) \leq Mg(x)$ for all possible values of x.
 
-For the first region ($-\infty, s-b$) $\pi(x)=0$ thus we can let $M_{1}$ be some arbitrarily small value greater than 0 say $\epsilon$ and this will satisfy $\pi(x) \leq M_{1}g(x)$ in the region specified.
+#### Region 1:
+For the first region ($-\infty, s-b$), $\pi(x)=0$ thus we can let $M_{1}$ be some arbitrarily small value greater than 0 say $\epsilon$.
+This will satisfy $\pi(x) \leq M_{1}g(x)$ in the region specified as $g(x) > 0$ for all values of $x$ in this region.
+
+#### Region 2:
 For the second region [$s-b,s$], $\pi(x)=\frac{1}{2b}$.
-Considering that $g(x)$ is an increasing function in this region (see plot 1) it will be at it's minimum value at $x=s-b$ and that the value of $\pi(x)$ is constant, setting $M_{2}=\frac{\pi(s-b)}{g(s-b)}=\frac{exp(\lambda|s-2b|)}{\lambda b}$ will satisfy the equation $\pi(x) \leq M_{2}g(x)$ for all x in the region.
+Considering that $g(x)$ is an increasing function in this region (see plot 1) it will be at it's minimum value at $x=s-b$ and that the value of $\pi(x)$ is constant.
+Thus, setting $M_{2}=\frac{\pi(s-b)}{g(s-b)}=\frac{exp(\lambda|s-2b|)}{\lambda b}$ will satisfy the equation $\pi(x) \leq M_{2}g(x)$ for all values of  $x$ in this region.
+
+#### Region 3:
 For the third region ($s,\infty$), $\pi(x)=\frac{as^{a}}{2x^{a+1}}$.
-I note that $\pi(x)$ is a decreasing function in this region (see plot 1) and initially $g(x)$ is a smaller but increasing function (see plot 1), until it reaches its maximum (larger than $\pi(x)$ at this value of x) where it then decreases for all larger x but always is larger than $\pi(x)$.
-Thus setting $M_{3}=\frac{\pi(s)}{g(s)}=\frac{a}{\lambda s}exp(\lambda |s-b|)$ will satisfy the equation $\pi(x) \leq M_{3}g(x)$ for all x in this region.
+I note that $\pi(x)$ is a decreasing function in this region (see plot 1) and initially $g(x)$ is a smaller but increasing function (see plot 1), until it reaches its maximum value at which point it is larger than $\pi(x)$.
+It then decreases for all larger x but always is larger than $\pi(x)$.
+Thus, setting $M_{3}=\frac{\pi(s)}{g(s)}=\frac{a}{\lambda s}exp(\lambda |s-b|)$ will satisfy the equation $\pi(x) \leq M_{3}g(x)$ for all $x$ in this region.
 
+#### Calculating M and choosing lambda:
 Subbing in the values of a=2, b=3 and s=1.25 given in the question the value of M for $\lambda=1$ is $M=max(\epsilon,\frac{e^{4.75}}{3},1.6e^{1.75}) \approx 38.53$ and for $\lambda=0.5$ is $M=max(\epsilon,\frac{e^{2.375}}{1.5},3.2e^{0.875})\approx 7.7$.
-Thus as M is smaller for $\lambda=0.5$.
-Thus, out of these 2 values for lambda 0.5 is my chosen value as it leads to an algorithm with a higher efficiency.
+Thus, as M is smaller for $\lambda=0.5$, 0.5 is my chosen value of lambda as it leads to an algorithm with a higher efficiency.
 
@@
 ## Part b:
-To implement A-R sampling using the proposal density with $\lambda=0.5$ to obtain samples from $\pi(x)$ I use the following algorithm:
+To implement A-R sampling using the proposal density with $\lambda=0.5$ to obtain samples from $\pi(x)$ with parameters $s$, $a$ and $b$ I use the following algorithm:
 
-To sample from $g(x)$ I use the inverse transformation technique by drawing u from a uniform distribution then if $u<0.5$ set $x=b+\frac{log(2u)}{\lambda}$ else set $x=b-\frac{2-2u}{\lambda}$.
-Then after obtaining x a draw from $g(x)$ perform A-R sampling by drawing y from a uniform(0,1) distribution and if $s < x$ then checking if $y \leq \frac{\pi(x)}{Mg(x)}=\frac{as^{a}}{2x^{a+1}}\frac{2}{M\lambda exp(-\lambda|x-b|)}$ else if $s-b <x$ then checking if $y \leq \frac{\pi(x)}{Mg(x)}=\frac{1}{2b}\frac{2}{M\lambda exp(-\lambda|x-b|)}$ if either of these checks are true then x is a draw from the target distribution $\pi(x)$ and we save this value, if these checks are not true or $x<s-b$ then x is not a draw from $\pi(x)$ and we throw away x.
-I repeat this process until the desired number of samples from $\pi(x)$ has been attained.
+#### Sampling from the proposal distribution:
+To sample from $g(x)$ I used the inverse transformation technique by drawing $u$ from a uniform distribution then if $u<0.5$ set $x=b+\frac{log(2u)}{\lambda}$ else set $x=b-\frac{2-2u}{\lambda}$.
+To obtain the value $x$, drawn from $g(x)$ the proposal distribution.
+
+#### Accept or reject:
+To either accept or reject the value $x$ I draw y from a Uniform(0,1) distribution and if:
+
+* $s < x$, check if $y \leq \frac{\pi(x)}{Mg(x)}=\frac{as^{a}}{2x^{a+1}}\frac{2}{M\lambda exp(-\lambda|x-b|)}$
+
+* $s-b <x$, check if $y \leq \frac{\pi(x)}{Mg(x)}=\frac{1}{2b}\frac{2}{M\lambda exp(-\lambda|x-b|)}$
+
+If either of these checks are true then $x$ is a draw from the target distribution $\pi(x)$ and the value is saved, if these checks are not true or $x<s-b$ then $x$ is not a draw from $\pi(x)$ and the value of $x$ is disregarded.
+Repeat this process until the desired number of samples from $\pi(x)$ has been attained.
 
@@
      xlab = "Values generated by AR sampling",
-     main = "Simulated and actual pdf of target distribution");
+     main = "Simulated and actual pdf of target distribution \n with AR generated sample plotted as a histogram");
 curve(target.dist(x,2,3,1.25),from=-4,to=20,n=1000,add=T, lwd= 1, lty= 1,col="red");
@@
 ```
+
+# Question 2
+
+## Part a
+For the continuous-time random walk $U=(U_{i})_{i \in \{1,...,d\}}$ conditioned on $U_{0}=u_{0}$:
+\begin{equation}
+U_{i} \sim \text{Normal}(\mu=U_{i-1},\sigma^{2}=t_{i}-t_{i-1}) \hspace{2em} \text{at time $t_{i}$}
+\end{equation}
+
+First by transforming $\boldsymbol{a} = (a_{1},...,a_{d})$ to the logit scale by setting $b_{i}=log(\frac{a_{i}}{1-a_{i}})$, $p_{tail}(\boldsymbol{a})=Pr(\boldsymbol{X}>\boldsymbol{a}|X_{0}=x_{0})  \equiv Pr(\boldsymbol{U}>\boldsymbol{b}|U_{0}=u_{0})=p_{tail}(\boldsymbol{b}))$, thus we can work with the sequence $U$ to find $p_{tail}(\boldsymbol{a})$.
+Now from equation (3):
+\begin{equation*}
+Pr(U_{i}>b_{i})=Pr((\sqrt{t_{i}-t_{i-1}})Z+u_{i-1}>b_{i}))=Pr\left(Z>\frac{b_{i}-u_{i-1}}{\sqrt{t_{i}-t_{i-1}}}\right) \hspace{2em}  \text{for} \hspace{1em} i \in \{0,...,d\}
+\end{equation*}
+For $Z$ the standard normal and $u_{i-1}$ a value drawn from $U_{i-1}$.
+
+Therefore, to not waste any sample values let $q_{i}$ the proposal distribution for the ith random variable in $U=(U_{i})_{i \in \{1,...,d\}}$ be an exponential distribution shifted by $s_{i}$ where:
+
+\begin{equation*}
+ s_{i}=\frac{b_{i}-u_{i-1}}{\sqrt{t_{i}-t_{i-1}}} 
+\end{equation*}
+
+As the samples $\boldsymbol{w}=\{w_{1},...,w_{d}\}$ drawn from these proposal distributions $\boldsymbol{q}=\{q_{1},....,q_{d}\}$ will satisfy $w_{i}>b_{i}$, $i \in \{1,...,d\}$ all values can be used to calculate $p_{tail}(\boldsymbol{b})=p_{tail}(\boldsymbol{a})$.
+
+To calculate $q_{i}$ however, we need to know the value $u_{i-1}$ (a value drawn from the random variable $U_{i-1}$) for each $i$ .
+Thus, each iteration will need to generate $\boldsymbol{u} = (u_{1},...,u_{d})$ a realization of the random walk sequence $U=(U_{i})_{i \in \{1,...,d\}}$ conditioned on $U_{0}=u_{0}$ using equation (3) with the initial value $u_{0}$ of the random walk and times $t_{i}$,  $i \in \{0,...,d\}$.
+Then storing these values to be accessed to generate a sample from the proposal distribution.
+
+## Part b
+
+#### Transforming inputs:
+The algorithm will start by transforming inputs from percentages to proportions by dividing values by 100 then applying the logit transform to $\boldsymbol{a}$ and calling the new vector of values $\boldsymbol{b}=log(\frac{\boldsymbol{a}}{1-\boldsymbol{a}})$ (as mentioned in part a)
+
+#### Drawing values from proposal distributions:
+Next for each iteration j we must simulate the random walk $U$ using equation 3 and storing these realizations $\boldsymbol{u}$ as mentioned in part a.
+Using these values and the relevant proposal distribution $q_{i}$ (described in part a) generate $w_{i}$ for each step $i$ the random walk sequence $U$.
+
+#### Calculating weights:
+Next calculate weights $w_{i}$ using the formula $w_{i}=\frac{\pi_{i}(w_{i})}{Q_{i}(w_{i})})$ where $\pi_{i}$ is the pdf of random variable $U_{i}$ (defined by equation 3) and $Q_{i}$ is the density of the proposal distribution $q_{i}$ an exponential distribution with rate $\lambda$ and shifted by $s_{i}$ defined in part a.
+
+Then the combined weight used to calculate $p_{tail}(\boldsymbol{b})$ is the product of the weights $\boldsymbol{w}=(w_{1},...,w_{d})$.
+This defines the weight of $\boldsymbol{U}>\boldsymbol{b}$ using the fact that $w_{i}$ is weight of $Pr(U_{i}>b_{i}|U_{i-1}>b_{i-1})$, the markov property and product of conditional probability.
+
+#### Calculating probability:
+To calculate $p_{tail}(\boldsymbol{b})$, The previous 2 steps are performed n times thus we have n values of the combined weight (which I will denote $\hat{w}$) which we then use to calculate the importance sampling estimator defined as:
+
+\begin{equation*}
+  I_{n}^{IS} = \frac{1}{n}\sum_{i=1}^{n}\hat{w}(U)\psi(U) = \frac{1}{n}\sum_{i=1}^{n}\hat{w}(U) \hspace{1em} \text{as} \hspace{1em} \psi(U)=1
+\end{equation*}
+
+Thus the IS estimator of $p_{tail}(\boldsymbol{b})$ is the mean of the n combined weights and this as discussed in part a is the IS estimator of $p_{tail}(\boldsymbol{a})$ too.
+
+# Question 3
+
+## Part a
+
+#### Initialisation:
+At initialization set $\boldsymbol{P}_{\boldsymbol{T}}^{cur}$ to be the initial value of $\boldsymbol{P}_{\boldsymbol{T}}$ specified in the function.
+
+#### Drawing new proposal value:
+For each iteration j start by drawing $\boldsymbol{P}_{\boldsymbol{T}}^{new}$ the proposed value of $\boldsymbol{P}_{\boldsymbol{T}}$ from the proposal density by drawing each $p_{x}'$ with $x \in \{s,c,r\}$ from Dirichlet $(20p_{x}^{(cur)})$ and combining these to get $\boldsymbol{P}_{\boldsymbol{T}}^{new}=(p_{s}',p_{c}',p_{r}')^{T}$.
+
+#### Calculate ratio for posterior density (target distribution):
+Calculate the ratio of the posterior density using the new and current values of $\boldsymbol{P}_{\boldsymbol{T}}$. First by considering that:
+
+\begin{equation*}
+ p(\boldsymbol{P}_{\boldsymbol{T}}|X,\pi) \propto p(X|\boldsymbol{P}_{\boldsymbol{T}},\pi)
+\end{equation*}
+
+We get $\frac{p(\boldsymbol{P}_{\boldsymbol{T}}^{new}|\boldsymbol{X},\pi)}{p(\boldsymbol{P}_{\boldsymbol{T}}^{(cur)}|\boldsymbol{X},\pi)}=\frac{p(\boldsymbol{X}|\boldsymbol{P}_{\boldsymbol{T}}^{new},\pi)}{p(\boldsymbol{X}|\boldsymbol{P}_{\boldsymbol{T}}^{cur},\pi)}$, thus we can calculate this ratio of densities as we can calculate the RHS.
+
+To calculate the RHS using the Markov property and equation 8 and 9 in the coursework question sheet:
+
+\begin{equation*}
+\begin{aligned}
+Pr(\boldsymbol{X}=\boldsymbol{x}|\pi,\boldsymbol{P}_{\boldsymbol{T}}) = Pr(X_{1}=x_{1}|\pi,,\boldsymbol{P}_{\boldsymbol{T}})Pr(X_{2}=x_{2}|X_{1}=x_{1},,\boldsymbol{P}_{\boldsymbol{T}})\cdot \cdot \cdot Pr(X_{n}=x_{n}|X_{n-1}=x_{n-1},,\boldsymbol{P}_{\boldsymbol{T}}) = \\ \pi_{x_{1}}\prod_{i=2}^{n}p_{T}(X_{i}=x_{i}|X_{i-1}=x_{i-1})
+\end{aligned}
+\end{equation*}
+
+#### Calculate ratio for proposal density:
+To calculate the ratio of the proposal density $q$ of the current and new values of $\boldsymbol{P}_{\boldsymbol{T}}$.
+Calculate the probability density of q=Dirichlet($20 p^{cur}_{x}$) with $v= p^{cur}_{x}$ and $v= p^{new}_{x}$ for each of the rows of $\boldsymbol{P}_{\boldsymbol{T}}^{cur}$ and $\boldsymbol{P}_{\boldsymbol{T}}^{new}$.
+Then calculate the the probability density of $\boldsymbol{P}_{\boldsymbol{T}}^{cur}$ and $\boldsymbol{P}_{\boldsymbol{T}}^{new}$ by finding the joint probability density of the rows $p_{x}$ by multiplying the probability densities of each row together (as independent can do this).
+Thus the ratio of the proposal density is:
+
+\begin{equation*}
+ \frac{q(\boldsymbol{P}_{\boldsymbol{T}}^{cur})}{q(\boldsymbol{P}_{\boldsymbol{T}}^{new})} = \frac{\prod_{x=1}^{3}q(p_{x}^{cur})}{\prod_{x=1}^{3}q(p_{x}^{new})}  \hspace{1em} \text{with} \hspace{1em} x \in \{1=s,2=c,3=r\}
+\end{equation*}
+
+#### Accept or reject proposal:
+Calculate $\alpha = min \left( 1,\frac{p(\boldsymbol{P}_{\boldsymbol{T}}^{new}|X,\pi)}{p(\boldsymbol{P}_{\boldsymbol{T}}^{cur}|X,\pi)}\frac{q(\boldsymbol{P}_{\boldsymbol{T}}^{cur})}{q(\boldsymbol{P}_{\boldsymbol{T}}^{new})}\right)$ using the ratios calculated above and generate $u$ a value drawn from a Uniform(0,1) distribution and if $u<=\alpha$ set $\boldsymbol{P}_{\boldsymbol{T}}^{new}$ to be the jth value of $\boldsymbol{P}_{\boldsymbol{T}}$ and update $\boldsymbol{P}_{\boldsymbol{T}}^{cur}$ to be the value of $\boldsymbol{P}_{\boldsymbol{T}}^{new}$.
+Otherwise, set $\boldsymbol{P}_{\boldsymbol{T}}^{cur}$ to be jth value of $\boldsymbol{P}_{\boldsymbol{T}}$ and don't update the value of $\boldsymbol{P}_{\boldsymbol{T}}^{cur}$.
+
+#### Iterate:
+Repeat until the desired number of $\boldsymbol{P}_{\boldsymbol{T}}$ samples (specified number n plus b the specified burn in number), throw away the first b values of $\boldsymbol{P}_{\boldsymbol{T}}$.
+The remaining n values of $\boldsymbol{P}_{\boldsymbol{T}}$ is the estimated posterior distribution of $\boldsymbol{P}_{\boldsymbol{T}}$.

### 2022-12-28 draft_report.Rmd
Revision of report. Had to change alot in Q2 as was wrong what I was talking about ;P

@@
 
-For acceptance-rejection sampling (AR sampling) we need to find a constant $M>0$ such that $\pi(x) \leq Mg(x)$ for $\pi(x)$ the target distribution and $g(x)$ the proposal density defined in equations 1 and 2 above.
+For acceptance-rejection sampling (AR sampling) we need to find a constant $M>0$ such that $\pi(x) \leq Mg(x), \forall x \in \mathbb{R}$ for $\pi(x)$ the target distribution and $g(x)$ the proposal density defined in equations 1 and 2 above.
 This constant M determines the acceptance probability which is $\frac{1}{M}$ and hence the efficiency of the algorithm.
-For multiple possible values of $\lambda$ in $g(x)$ the value of $\lambda$ that results in a smaller constant M will have a higher acceptance probability and thus be our preferred value of $\lambda$ for the algorithm.
-To find M for each $\lambda$ value I split the real domain of x into 3 regions (see plot 1) that I call 1,2 and 3 which correspond to the different expressions of $\pi$ in different regions and study the relationship between $g(x)$ and $\pi(x)$ within these 3 regions to find $M_{1},M_{2},M_{3}$ which correspond to the smallest constant such that $\pi(x) \leq M_{i}g(x)$ for the ith region.
-Then set $M=max(M_{1},M_{2},M_{3})$ as this is the smallest constant that satisfies $\pi(x) \leq Mg(x)$ for all possible values of x.
+For multiple possible values of $\lambda$ in $g(x)$ the value of $\lambda$ that results in a smaller constant M will have a higher acceptance probability and thus will be our preferred value of $\lambda$ for the algorithm.
+To find M for each $\lambda$ value I split the real domain of $x$ into 3 regions (see plot 1) that I call 1,2 and 3 which correspond to the different expressions of $\pi(x)$ depending on the value of $x$.
+Then study the relationship between $g(x)$ and $\pi(x)$ within these 3 regions to find $M_{1},M_{2},M_{3}$ which correspond to the smallest constant such that $\pi(x) \leq M_{i}g(x)$ for the ith region.
+Then set $M=max(M_{1},M_{2},M_{3})$ as this is the smallest constant that satisfies $\pi(x) \leq Mg(x)$ for all possible values of $x \in \mathbb{R}$.
 
@@
 For the second region [$s-b,s$], $\pi(x)=\frac{1}{2b}$.
-Considering that $g(x)$ is an increasing function in this region (see plot 1) it will be at it's minimum value at $x=s-b$ and that the value of $\pi(x)$ is constant.
+Considering that $g(x)$ is an increasing function in this region (see plot 1) it will be at it's minimum value at $x=s-b$ and the value of $\pi(x)$ is constant.
 Thus, setting $M_{2}=\frac{\pi(s-b)}{g(s-b)}=\frac{exp(\lambda|s-2b|)}{\lambda b}$ will satisfy the equation $\pi(x) \leq M_{2}g(x)$ for all values of  $x$ in this region.
@@
 I note that $\pi(x)$ is a decreasing function in this region (see plot 1) and initially $g(x)$ is a smaller but increasing function (see plot 1), until it reaches its maximum value at which point it is larger than $\pi(x)$.
-It then decreases for all larger x but always is larger than $\pi(x)$.
+It then becomes a decreasing function of $x$ but is always larger than $\pi(x)$.
 Thus, setting $M_{3}=\frac{\pi(s)}{g(s)}=\frac{a}{\lambda s}exp(\lambda |s-b|)$ will satisfy the equation $\pi(x) \leq M_{3}g(x)$ for all $x$ in this region.
@@
 Subbing in the values of a=2, b=3 and s=1.25 given in the question the value of M for $\lambda=1$ is $M=max(\epsilon,\frac{e^{4.75}}{3},1.6e^{1.75}) \approx 38.53$ and for $\lambda=0.5$ is $M=max(\epsilon,\frac{e^{2.375}}{1.5},3.2e^{0.875})\approx 7.7$.
-Thus, as M is smaller for $\lambda=0.5$, 0.5 is my chosen value of lambda as it leads to an algorithm with a higher efficiency.
+Thus, as M is smaller for $\lambda=0.5$, 0.5 is my chosen value of lambda as it leads to an algorithm with a higher acceptance probability.
 
@@
      main="Plot 1: Comparing target and proposal pdf's",
-     sub="Note: used a=2,b=3,s=1.25 with the green line seperating the regions of the graph")
+     sub="Note: used a=2, b=3, s=1.25 with the green line seperating the regions of the graph")
 curve(gx.1(x),add=T,col="red")
@@
 ## Part b:
-To implement A-R sampling using the proposal density with $\lambda=0.5$ to obtain samples from $\pi(x)$ with parameters $s$, $a$ and $b$ I use the following algorithm:
+To implement A-R sampling using the proposal density with $\lambda=0.5$ and thus $M=3.2e^{0.875}$ to obtain samples from $\pi(x)$ with parameters $s$, $a$ and $b$ I use the following algorithm:
 
 #### Sampling from the proposal distribution:
-To sample from $g(x)$ I used the inverse transformation technique by drawing $u$ from a uniform distribution then if $u<0.5$ set $x=b+\frac{log(2u)}{\lambda}$ else set $x=b-\frac{2-2u}{\lambda}$.
+To sample from $g(x)$, use the inverse transformation technique by drawing $u$ from a Uniform(0,1) distribution then if $u<0.5$ set $x=b+\frac{log(2u)}{\lambda}$, else set $x=b-\frac{2-2u}{\lambda}$.
 To obtain the value $x$, drawn from $g(x)$ the proposal distribution.
@@
 #### Accept or reject:
-To either accept or reject the value $x$ I draw y from a Uniform(0,1) distribution and if:
+To either accept or reject the value $x$, draw y from a Uniform(0,1) distribution and if:
 
@@
 
-First by transforming $\boldsymbol{a} = (a_{1},...,a_{d})$ to the logit scale by setting $b_{i}=log(\frac{a_{i}}{1-a_{i}})$, $p_{tail}(\boldsymbol{a})=Pr(\boldsymbol{X}>\boldsymbol{a}|X_{0}=x_{0})  \equiv Pr(\boldsymbol{U}>\boldsymbol{b}|U_{0}=u_{0})=p_{tail}(\boldsymbol{b}))$, thus we can work with the sequence $U$ to find $p_{tail}(\boldsymbol{a})$.
+First, by transforming $\boldsymbol{a} = (a_{1},...,a_{d})$ to the logit scale by setting $b_{i}=log(\frac{a_{i}}{1-a_{i}})$, $p_{tail}(\boldsymbol{a})=Pr(\boldsymbol{X}>\boldsymbol{a}|X_{0}=x_{0})  \equiv Pr(\boldsymbol{U}>\boldsymbol{b}|U_{0}=u_{0})=p_{tail}(\boldsymbol{b})$, thus we can work with the sequence $U$ to find $p_{tail}(\boldsymbol{a})$.
 Now from equation (3):
@@
 For $Z$ the standard normal and $u_{i-1}$ a value drawn from $U_{i-1}$.
+Therefore, to find $Pr(U_{i}>b_{i})$ we can find the probability a standard normal distribution is greater than $\frac{b_{i}-u_{i-1}}{\sqrt{t_{i}-t_{i-1}}}$.
+Therefore, letting $\hat{p}_{t}^{(i)}$ be the importance sampling estimator of $p_{t}^{(i)}=Pr(U_{i}>b_{i})$ we can use the following equation:
 
-Therefore, to not waste any sample values let $q_{i}$ the proposal distribution for the ith random variable in $U=(U_{i})_{i \in \{1,...,d\}}$ be an exponential distribution shifted by $s_{i}$ where:
+\begin{equation*}
+ \hat{p}_{t}^{(i)} = \frac{1}{n}\sum_{j=1}^{n}\frac{\phi(V_{j})}{q_{i}(V_{j})}I\left(V_{j}>\frac{b_{i}-u_{i-1}}{\sqrt{t_{i}-t_{i-1}}} \right)
+\end{equation*}
+
+With $V_{j}, j \in \{1,..,n\}$ iid random variables from the proposal distribution $q_{i}$, $\phi$ the density of a standard normal distribution and $I$ the indicator function.
+Therefore, to not waste any sample values let $q_{i}$ be an exponential distribution shifted by $s_{i}$ where:
 
@@
 
-As the samples $\boldsymbol{w}=\{w_{1},...,w_{d}\}$ drawn from these proposal distributions $\boldsymbol{q}=\{q_{1},....,q_{d}\}$ will satisfy $w_{i}>b_{i}$, $i \in \{1,...,d\}$ all values can be used to calculate $p_{tail}(\boldsymbol{b})=p_{tail}(\boldsymbol{a})$.
+As all sample values drawn from these proposal distributions $\boldsymbol{q}=\{q_{1},....,q_{d}\}$ will satisfy $v_{i}>\frac{b_{i}-u_{i-1}}{\sqrt{t_{i}-t_{i-1}}}$ for $i \in \{1,...,d\}$.
+Thus, all values generated from these proposal distributions can be used to calculate $\hat{p}_{t}^{(i)}$ for $i \in \{1,..,d\}$ (as the indicator function is 1 for all samples drawn to calculate each $\hat{p}_{t}^{(i)}$).
 
 To calculate $q_{i}$ however, we need to know the value $u_{i-1}$ (a value drawn from the random variable $U_{i-1}$) for each $i$ .
-Thus, each iteration will need to generate $\boldsymbol{u} = (u_{1},...,u_{d})$ a realization of the random walk sequence $U=(U_{i})_{i \in \{1,...,d\}}$ conditioned on $U_{0}=u_{0}$ using equation (3) with the initial value $u_{0}$ of the random walk and times $t_{i}$,  $i \in \{0,...,d\}$.
+Thus, each iteration will need to generate $\boldsymbol{u} = (u_{1},...,u_{d-1})$ a realization of the random walk sequence $U=(U_{i})_{i \in \{1,...,d-1\}}$ conditioned on $U_{0}=u_{0}$ using equation (3) with the initial value $u_{0}$ of the random walk and times $t_{i}$,  $i \in \{0,...,d\}$.
 Then storing these values to be accessed to generate a sample from the proposal distribution.
+
+Thus, using $\boldsymbol{q}$ we can generate generate the sample vector $\boldsymbol{v}=\{v_{1},....,v_{d}\}$ where each $v_{i}$ satisfies $v_{i}>\frac{b_{i}-u_{i-1}}{\sqrt{t_{i}-t_{i-1}}}$.
+Thus, as $\prod_{i=1}^{d}\hat{p}_{t}^{(i)}=p_{tail}(\boldsymbol{b})$ each vector is used to calculate $p_{tail}(\boldsymbol{b})=p_{tail}(\boldsymbol{a})$.
 
@@
 #### Transforming inputs:
-The algorithm will start by transforming inputs from percentages to proportions by dividing values by 100 then applying the logit transform to $\boldsymbol{a}$ and calling the new vector of values $\boldsymbol{b}=log(\frac{\boldsymbol{a}}{1-\boldsymbol{a}})$ (as mentioned in part a)
+The algorithm will start by transforming inputs from percentages to proportions by dividing the values by 100 then applying the logit transform to $\boldsymbol{a}$ and calling the new vector of values $\boldsymbol{b}=log(\frac{\boldsymbol{a}}{1-\boldsymbol{a}})$ (as mentioned in part a)
 
 #### Drawing values from proposal distributions:
-Next for each iteration j we must simulate the random walk $U$ using equation 3 and storing these realizations $\boldsymbol{u}$ as mentioned in part a.
-Using these values and the relevant proposal distribution $q_{i}$ (described in part a) generate $w_{i}$ for each step $i$ the random walk sequence $U$.
+Next for each iteration j we must simulate the random walk $U$ using equation 3 and storing the realization $\boldsymbol{u}$ as mentioned in part a.
+Using these values and the relevant proposal distribution $q_{i}$ (described in part a) to generate $v_{i}$ a sample from $q_{i}$ for each step $i$ of the random walk sequence $U$.
 
 #### Calculating weights:
-Next calculate weights $w_{i}$ using the formula $w_{i}=\frac{\pi_{i}(w_{i})}{Q_{i}(w_{i})})$ where $\pi_{i}$ is the pdf of random variable $U_{i}$ (defined by equation 3) and $Q_{i}$ is the density of the proposal distribution $q_{i}$ an exponential distribution with rate $\lambda$ and shifted by $s_{i}$ defined in part a.
+Next calculate weights $w_{i}$ of each $v_{i}$ using the formula $w_{i}=\frac{\phi(v_{i})}{q_{i}(v_{i})})$ where $\phi$ is the demnsity of the standard normal and $q_{i}$ is the density of the proposal distribution, an exponential distribution with rate $\lambda$ and shifted by $s_{i}$ as defined in part a.
 
-Then the combined weight used to calculate $p_{tail}(\boldsymbol{b})$ is the product of the weights $\boldsymbol{w}=(w_{1},...,w_{d})$.
-This defines the weight of $\boldsymbol{U}>\boldsymbol{b}$ using the fact that $w_{i}$ is weight of $Pr(U_{i}>b_{i}|U_{i-1}>b_{i-1})$, the markov property and product of conditional probability.
+Then the combined weight used to calculate $p_{tail}(\boldsymbol{b})$ is the product of the weights $\prod_{i=1}^{n}w_{i}$.
+This is the weight used to calculate $Pr(\boldsymbol{U}>\boldsymbol{b}|U_{0}=u_{0})$ (using the fact that $w_{i}$ is weight of $Pr(U_{i}>b_{i}|U_{0}=u_{0})$).
 
 #### Calculating probability:
-To calculate $p_{tail}(\boldsymbol{b})$, The previous 2 steps are performed n times thus we have n values of the combined weight (which I will denote $\hat{w}$) which we then use to calculate the importance sampling estimator defined as:
+To calculate $p_{tail}(\boldsymbol{b})$, The previous 2 steps are performed n times thus we have n values of the combined weight (which I will denote $\hat{w}$) which we then use to calculate the importance sampling (IS) estimator defined as:
 
@@
 
+\newpage
 # Question 3
@@
 #### Drawing new proposal value:
-For each iteration j start by drawing $\boldsymbol{P}_{\boldsymbol{T}}^{new}$ the proposed value of $\boldsymbol{P}_{\boldsymbol{T}}$ from the proposal density by drawing each $p_{x}'$ with $x \in \{s,c,r\}$ from Dirichlet $(20p_{x}^{(cur)})$ and combining these to get $\boldsymbol{P}_{\boldsymbol{T}}^{new}=(p_{s}',p_{c}',p_{r}')^{T}$.
+For each iteration j start by drawing $\boldsymbol{P}_{\boldsymbol{T}}^{new}$ the proposed value of $\boldsymbol{P}_{\boldsymbol{T}}$ from the proposal density by drawing each $p_{x}'$ with $x \in \{s,c,r\}$ from a Dirichlet $(20p_{x}^{(cur)})$ distribution and combining these to get $\boldsymbol{P}_{\boldsymbol{T}}^{new}=(p_{s}',p_{c}',p_{r}')^{T}$.
 
-#### Calculate ratio for posterior density (target distribution):
+#### Calculate the ratio for the posterior density (target distribution):
 Calculate the ratio of the posterior density using the new and current values of $\boldsymbol{P}_{\boldsymbol{T}}$. First by considering that:
@@
 
-To calculate the RHS using the Markov property and equation 8 and 9 in the coursework question sheet:
+To calculate the RHS we use the Markov property and equation 8 and 9 in the coursework question sheet to get following equation:
 
@@
 
-#### Calculate ratio for proposal density:
+Therefore, subbing in $\boldsymbol{P}_{\boldsymbol{T}}^{new}$ and $\boldsymbol{P}_{\boldsymbol{T}}^{cur}$ in the above equation and then calculating the ratio of results we can calculate the ratio of the posterior density.
+
+#### Calculate the ratio for the proposal density:
 To calculate the ratio of the proposal density $q$ of the current and new values of $\boldsymbol{P}_{\boldsymbol{T}}$.
-Calculate the probability density of q=Dirichlet($20 p^{cur}_{x}$) with $v= p^{cur}_{x}$ and $v= p^{new}_{x}$ for each of the rows of $\boldsymbol{P}_{\boldsymbol{T}}^{cur}$ and $\boldsymbol{P}_{\boldsymbol{T}}^{new}$.
+Calculate the probability density of a Dirichlet($20 p^{cur}_{x}$) distribution (the proposal distribution) which I will denote $f(v)$ with $v$ equal to either $p^{cur}_{x}$ or $v= p^{new}_{x}$ for each of the rows of $\boldsymbol{P}_{\boldsymbol{T}}^{cur}$ or $\boldsymbol{P}_{\boldsymbol{T}}^{new}$ respectively.
 Then calculate the the probability density of $\boldsymbol{P}_{\boldsymbol{T}}^{cur}$ and $\boldsymbol{P}_{\boldsymbol{T}}^{new}$ by finding the joint probability density of the rows $p_{x}$ by multiplying the probability densities of each row together (as independent can do this).
@@
 \begin{equation*}
- \frac{q(\boldsymbol{P}_{\boldsymbol{T}}^{cur})}{q(\boldsymbol{P}_{\boldsymbol{T}}^{new})} = \frac{\prod_{x=1}^{3}q(p_{x}^{cur})}{\prod_{x=1}^{3}q(p_{x}^{new})}  \hspace{1em} \text{with} \hspace{1em} x \in \{1=s,2=c,3=r\}
+ \frac{f(\boldsymbol{P}_{\boldsymbol{T}}^{cur})}{f(\boldsymbol{P}_{\boldsymbol{T}}^{new})} = \frac{\prod_{x=1}^{3}f(p_{x}^{cur})}{\prod_{x=1}^{3}f(p_{x}^{new})}  \hspace{1em} \text{with} \hspace{1em} x \in \{1=s,2=c,3=r\}
 \end{equation*}
@@
 #### Accept or reject proposal:
-Calculate $\alpha = min \left( 1,\frac{p(\boldsymbol{P}_{\boldsymbol{T}}^{new}|X,\pi)}{p(\boldsymbol{P}_{\boldsymbol{T}}^{cur}|X,\pi)}\frac{q(\boldsymbol{P}_{\boldsymbol{T}}^{cur})}{q(\boldsymbol{P}_{\boldsymbol{T}}^{new})}\right)$ using the ratios calculated above and generate $u$ a value drawn from a Uniform(0,1) distribution and if $u<=\alpha$ set $\boldsymbol{P}_{\boldsymbol{T}}^{new}$ to be the jth value of $\boldsymbol{P}_{\boldsymbol{T}}$ and update $\boldsymbol{P}_{\boldsymbol{T}}^{cur}$ to be the value of $\boldsymbol{P}_{\boldsymbol{T}}^{new}$.
+Calculate $\alpha = min \left( 1,\frac{p(\boldsymbol{P}_{\boldsymbol{T}}^{new}|X,\pi)}{p(\boldsymbol{P}_{\boldsymbol{T}}^{cur}|X,\pi)}\frac{f(\boldsymbol{P}_{\boldsymbol{T}}^{cur})}{f(\boldsymbol{P}_{\boldsymbol{T}}^{new})}\right)$ using the ratios calculated above and generate $u$ a value drawn from a Uniform(0,1) distribution and if $u<=\alpha$ set $\boldsymbol{P}_{\boldsymbol{T}}^{new}$ to be the jth value of $\boldsymbol{P}_{\boldsymbol{T}}$ and update $\boldsymbol{P}_{\boldsymbol{T}}^{cur}$ to be the value of $\boldsymbol{P}_{\boldsymbol{T}}^{new}$.
 Otherwise, set $\boldsymbol{P}_{\boldsymbol{T}}^{cur}$ to be jth value of $\boldsymbol{P}_{\boldsymbol{T}}$ and don't update the value of $\boldsymbol{P}_{\boldsymbol{T}}^{cur}$.
@@
 #### Iterate:
-Repeat until the desired number of $\boldsymbol{P}_{\boldsymbol{T}}$ samples (specified number n plus b the specified burn in number), throw away the first b values of $\boldsymbol{P}_{\boldsymbol{T}}$.
+Repeat until we have attained n+b+1 sample values of $\boldsymbol{P}_{\boldsymbol{T}}$ where n is the number of desired samples and b is the specified burn in number.
+Then throw away the first b+1 values of $\boldsymbol{P}_{\boldsymbol{T}}$.
 The remaining n values of $\boldsymbol{P}_{\boldsymbol{T}}$ is the estimated posterior distribution of $\boldsymbol{P}_{\boldsymbol{T}}$.

### 2022-12-28 Final.code.R

@@ -43,7 +43,7 @@ plot(x=x,y=target.dist(x),
      type="l",
      ylab = "density",
      main="Plot 1: Comparing target and proposal pdf's",
-     sub="Note: used a=2,b=3,s=1.25 with the green line seperating the regions of the graph")
+     sub="Note: used a=2, b=3, s=1.25 with the green line seperating the regions of the graph")
 curve(gx.1(x),add=T,col="red")
 curve(gx.2,add=T,col="blue")
 abline(v=-1.75,lty=2,col="green")

### 2022-12-30 draft_report.Rmd
MOre big changes to Q2 in 2nd review of report

@@
 
-For acceptance-rejection sampling (AR sampling) we need to find a constant $M>0$ such that $\pi(x) \leq Mg(x), \forall x \in \mathbb{R}$ for $\pi(x)$ the target distribution and $g(x)$ the proposal density defined in equations 1 and 2 above.
+For acceptance-rejection sampling (A-R sampling) we need to find a constant $M>0$ such that $\pi(x) \leq Mg(x), \forall x \in \mathbb{R}$.
+For $\pi(x)$ the target density (the density from which we want to generate a sample of random values) and $g(x)$ the proposal density defined in equations 1 and 2 above respectively.
 This constant M determines the acceptance probability which is $\frac{1}{M}$ and hence the efficiency of the algorithm.
 For multiple possible values of $\lambda$ in $g(x)$ the value of $\lambda$ that results in a smaller constant M will have a higher acceptance probability and thus will be our preferred value of $\lambda$ for the algorithm.
-To find M for each $\lambda$ value I split the real domain of $x$ into 3 regions (see plot 1) that I call 1,2 and 3 which correspond to the different expressions of $\pi(x)$ depending on the value of $x$.
-Then study the relationship between $g(x)$ and $\pi(x)$ within these 3 regions to find $M_{1},M_{2},M_{3}$ which correspond to the smallest constant such that $\pi(x) \leq M_{i}g(x)$ for the ith region.
+To find M for each of the proposed $\lambda$ values I split the real domain of $x$ into 3 regions (see plot 1) that I call 1,2 and 3 which correspond to the different expressions of $\pi(x)$ depending on the value of $x$.
+Then study the relationship between $g(x)$ and $\pi(x)$ within these 3 regions to find $M_{1},M_{2},M_{3}$ which correspond to the smallest constant such that $\pi(x) \leq M_{i}g(x)$ for all possibles values of x in the ith region.
 Then set $M=max(M_{1},M_{2},M_{3})$ as this is the smallest constant that satisfies $\pi(x) \leq Mg(x)$ for all possible values of $x \in \mathbb{R}$.
@@
 #### Calculating M and choosing lambda:
-Subbing in the values of a=2, b=3 and s=1.25 given in the question the value of M for $\lambda=1$ is $M=max(\epsilon,\frac{e^{4.75}}{3},1.6e^{1.75}) \approx 38.53$ and for $\lambda=0.5$ is $M=max(\epsilon,\frac{e^{2.375}}{1.5},3.2e^{0.875})\approx 7.7$.
+Subbing in the values of a=2, b=3 and s=1.25 given in the question, the value of M for $\lambda=1$ is $M=max(\epsilon,\frac{e^{4.75}}{3},1.6e^{1.75}) \approx 38.53$ and for $\lambda=0.5$ is $M=max(\epsilon,\frac{e^{2.375}}{1.5},3.2e^{0.875})\approx 7.7$.
 Thus, as M is smaller for $\lambda=0.5$, 0.5 is my chosen value of lambda as it leads to an algorithm with a higher acceptance probability.
@@
 curve(gx.1(x),add=T,col="red")
-curve(gx.2,add=T,col="blue")
+curve(gx.2(x),add=T,col="blue")
 abline(v=-1.75,lty=2,col="green")
@@
 
-## Part a
+## Part a:
 For the continuous-time random walk $U=(U_{i})_{i \in \{1,...,d\}}$ conditioned on $U_{0}=u_{0}$:
@@
 
-First, by transforming $\boldsymbol{a} = (a_{1},...,a_{d})$ to the logit scale by setting $b_{i}=log(\frac{a_{i}}{1-a_{i}})$, $p_{tail}(\boldsymbol{a})=Pr(\boldsymbol{X}>\boldsymbol{a}|X_{0}=x_{0})  \equiv Pr(\boldsymbol{U}>\boldsymbol{b}|U_{0}=u_{0})=p_{tail}(\boldsymbol{b})$, thus we can work with the sequence $U$ to find $p_{tail}(\boldsymbol{a})$.
+First, by transforming $\boldsymbol{a} = (a_{1},...,a_{d})$ to the logit scale by setting $b_{i}=log(\frac{a_{i}}{1-a_{i}})$, $p_{tail}(\boldsymbol{a})=Pr(\boldsymbol{X}>\boldsymbol{a}|X_{0}=x_{0})  \equiv Pr(U>\boldsymbol{b}) := p_{tail}(\boldsymbol{b})$ as $U_{i}=logit(X_{i})$ and the sequence U is already conditioned on $U_{0}=u_{0}$.
+Thus, we can work with the sequence $U$ and $\boldsymbol{b}$ to find $p_{tail}(\boldsymbol{a})$.
 Now from equation (3):
 \begin{equation*}
-Pr(U_{i}>b_{i})=Pr((\sqrt{t_{i}-t_{i-1}})Z+u_{i-1}>b_{i}))=Pr\left(Z>\frac{b_{i}-u_{i-1}}{\sqrt{t_{i}-t_{i-1}}}\right) \hspace{2em}  \text{for} \hspace{1em} i \in \{0,...,d\}
+Pr(U_{i}>b_{i})=Pr((\sqrt{t_{i}-t_{i-1}})Z+u_{i-1}>b_{i})=Pr\left(Z>\frac{b_{i}-u_{i-1}}{\sqrt{t_{i}-t_{i-1}}}\right) \hspace{2em}  \text{for} \hspace{1em} i \in \{1,...,d\}
 \end{equation*}
 For $Z$ the standard normal and $u_{i-1}$ a value drawn from $U_{i-1}$.
-Therefore, to find $Pr(U_{i}>b_{i})$ we can find the probability a standard normal distribution is greater than $\frac{b_{i}-u_{i-1}}{\sqrt{t_{i}-t_{i-1}}}$.
-Therefore, letting $\hat{p}_{t}^{(i)}$ be the importance sampling estimator of $p_{t}^{(i)}=Pr(U_{i}>b_{i})$ we can use the following equation:
+Therefore, to find $Pr(U_{i}>b_{i})$ we can find the probability a standard normal distribution is greater than $s_{i} := \frac{b_{i}-u_{i-1}}{\sqrt{t_{i}-t_{i-1}}}$.
+Therefore:
 
 \begin{equation*}
- \hat{p}_{t}^{(i)} = \frac{1}{n}\sum_{j=1}^{n}\frac{\phi(V_{j})}{q_{i}(V_{j})}I\left(V_{j}>\frac{b_{i}-u_{i-1}}{\sqrt{t_{i}-t_{i-1}}} \right)
+Pr(U>\boldsymbol{b})=\prod_{i=1}^{d}Pr\left(Z>\frac{b_{i}-u_{i-1}}{\sqrt{t_{i}-t_{i-1}}}\right)=\prod_{i=1}^{d}\int_{s_{i}}^{\infty}\phi(x)dx=\prod_{i=1}^{d}\int_{-\infty}^{\infty}\frac{\phi(x)}{q_{i}(x)}q_{i}(x)I(x>s_{i})dx = E\left[\prod_{i=1}^{d} \frac{\phi(V_{i})}{q_{i}(V_{i})}I(V_{i}>s_{i})\right]
 \end{equation*}
 
-With $V_{j}, j \in \{1,..,n\}$ iid random variables from the proposal distribution $q_{i}$, $\phi$ the density of a standard normal distribution and $I$ the indicator function.
-Therefore, to not waste any sample values let $q_{i}$ be an exponential distribution shifted by $s_{i}$ where:
+With $V_{i}, i \in \{1,..,d\}$ i.i.d.
+random variables from the proposal distribution, $q_{i}$ the density of the proposal distribution, $\phi$ the density of a standard normal distribution and $I$ the indicator function.
+
+Now using the above equation we get the importance sampling estimator for $p_{tail}(\boldsymbol{b})$ to be:
+
+\begin{equation}
+ I_{n}^{IS}=\frac{1}{n}\sum_{j=1}^{n}\left[ \prod_{i=1}^{d}w_{i}^{(j)}(V_{i}^{(j)})  I(V_{i}^{(j)}>s_{i}^{(j)}) \right]
+\end{equation}
+
+Where $w_{i}^{(j)}$ is $\frac{\phi(V_{i})}{q_{i}(V_{i})}$ on the jth iteration and the superscript $(j)$ denoting the jth sampling.
+Therefore, to not waste any sample values let the proposal distributions be exponential distributions shifted by $s_{i}$ as all sample values (denoted $v_{i}$) drawn from these proposal distributions will satisfy $v_{i}>s_{i}$ for $i \in \{1,...,d\}$.
+Thus, all values generated from these proposal distributions can be used to calculate $I_{n}^{IS}$ from equation (4) as the indicator function is 1 for each $v_{i}$ drawn from each $V_{i}$.
+
+To calculate $q_{i}$ however, we need to know the value $u_{i-1}$ (a value drawn from the random variable $U_{i-1}$) for each $i$.
+Thus, each iteration $j$ which generates a sample vector $\boldsymbol{v}$ will need to generate $\boldsymbol{u} = (u_{1},...,u_{d-1})$ a realization of the random walk sequence $U=(U_{i})_{i \in \{1,...,d-1\}}$ conditioned on $U_{0}=u_{0}$.
+This can be done by using equation (3) to generate a random walk when provided with an initial value $u_{0}$ of the random walk and times $t_{i}$,  $i \in \{0,...,d\}$.
+Then storing these values they can be accessed to generate $\boldsymbol{v}$ using the proposal distributions.
+
+Thus, using $d$ exponential distributions shifted by $s_{i}$ for $i \in \{1,..,d\}$ respectively we can generate the sample vector $\boldsymbol{v}>\boldsymbol{s}$ such that using equation (4):
 
 \begin{equation*}
- s_{i}=\frac{b_{i}-u_{i-1}}{\sqrt{t_{i}-t_{i-1}}} 
+ \hat{p}_{tail}(\boldsymbol{a})=\frac{1}{n}\sum_{j=1}^{n}\left[ \prod_{i=1}^{d}w_{i}^{(j)}(v_{i}^{(j)}) \right]
 \end{equation*}
 
-As all sample values drawn from these proposal distributions $\boldsymbol{q}=\{q_{1},....,q_{d}\}$ will satisfy $v_{i}>\frac{b_{i}-u_{i-1}}{\sqrt{t_{i}-t_{i-1}}}$ for $i \in \{1,...,d\}$.
-Thus, all values generated from these proposal distributions can be used to calculate $\hat{p}_{t}^{(i)}$ for $i \in \{1,..,d\}$ (as the indicator function is 1 for all samples drawn to calculate each $\hat{p}_{t}^{(i)}$).
-
-To calculate $q_{i}$ however, we need to know the value $u_{i-1}$ (a value drawn from the random variable $U_{i-1}$) for each $i$ .
-Thus, each iteration will need to generate $\boldsymbol{u} = (u_{1},...,u_{d-1})$ a realization of the random walk sequence $U=(U_{i})_{i \in \{1,...,d-1\}}$ conditioned on $U_{0}=u_{0}$ using equation (3) with the initial value $u_{0}$ of the random walk and times $t_{i}$,  $i \in \{0,...,d\}$.
-Then storing these values to be accessed to generate a sample from the proposal distribution.
-
-Thus, using $\boldsymbol{q}$ we can generate generate the sample vector $\boldsymbol{v}=\{v_{1},....,v_{d}\}$ where each $v_{i}$ satisfies $v_{i}>\frac{b_{i}-u_{i-1}}{\sqrt{t_{i}-t_{i-1}}}$.
-Thus, as $\prod_{i=1}^{d}\hat{p}_{t}^{(i)}=p_{tail}(\boldsymbol{b})$ each vector is used to calculate $p_{tail}(\boldsymbol{b})=p_{tail}(\boldsymbol{a})$.
+Where $\hat{p}_{tail}(\boldsymbol{a})$ is the IS estimate of ${p}_{tail}(\boldsymbol{a})$. Thus, each vector is used to calculate the IS estimate of $p_{tail}(\boldsymbol{a})$.
 
@@
 #### Drawing values from proposal distributions:
-Next for each iteration j we must simulate the random walk $U$ using equation 3 and storing the realization $\boldsymbol{u}$ as mentioned in part a.
-Using these values and the relevant proposal distribution $q_{i}$ (described in part a) to generate $v_{i}$ a sample from $q_{i}$ for each step $i$ of the random walk sequence $U$.
+Next for each iteration j we must simulate the random walk $U$ using equation (3), the initial value $u_{0}$ and times $t_{i}, i \in \{0,..,d\}$ and storing the values of the random walk $\boldsymbol{u}$ as mentioned in part a.
+Using these values and the relevant proposal distribution of an exponential distribution shifted by $s_{i}$ (described in part a) to generate $v_{i}, i \in \{1,..,d\}$.
 
 #### Calculating weights:
-Next calculate weights $w_{i}$ of each $v_{i}$ using the formula $w_{i}=\frac{\phi(v_{i})}{q_{i}(v_{i})})$ where $\phi$ is the demnsity of the standard normal and $q_{i}$ is the density of the proposal distribution, an exponential distribution with rate $\lambda$ and shifted by $s_{i}$ as defined in part a.
+Next calculate weights $w_{i}$ of each $v_{i}$ using the formula $w_{i}=\frac{\phi(v_{i})}{q_{i}(v_{i})}$ where $\phi$ is the density of the standard normal and $q_{i}$ is the density of the proposal distribution, an exponential distribution with rate $\lambda$ and shifted by $s_{i}$ as defined in part a.
 
-Then the combined weight used to calculate $p_{tail}(\boldsymbol{b})$ is the product of the weights $\prod_{i=1}^{n}w_{i}$.
-This is the weight used to calculate $Pr(\boldsymbol{U}>\boldsymbol{b}|U_{0}=u_{0})$ (using the fact that $w_{i}$ is weight of $Pr(U_{i}>b_{i}|U_{0}=u_{0})$).
+Then calculate the product of the weights $\prod_{i=1}^{n}w_{i}$ and save this as the jth value in a vector.
 
 #### Calculating probability:
-To calculate $p_{tail}(\boldsymbol{b})$, The previous 2 steps are performed n times thus we have n values of the combined weight (which I will denote $\hat{w}$) which we then use to calculate the importance sampling (IS) estimator defined as:
+To calculate $p_{tail}(\boldsymbol{b})$, the previous 2 steps are performed n times thus we have n values of the product of the weights (which I will denote $\hat{w}^{(j)}$ where the value of $j$ denotes the jth iteration) which we then use to calculate the importance sampling (IS) estimator using equation (4) in part a:
 
 \begin{equation*}
-  I_{n}^{IS} = \frac{1}{n}\sum_{i=1}^{n}\hat{w}(U)\psi(U) = \frac{1}{n}\sum_{i=1}^{n}\hat{w}(U) \hspace{1em} \text{as} \hspace{1em} \psi(U)=1
+  \hat{p}_{tail}(\boldsymbol{a}) = \frac{1}{n}\sum_{j=1}^{n}\hat{w}^{(j)}
 \end{equation*}
-
-Thus the IS estimator of $p_{tail}(\boldsymbol{b})$ is the mean of the n combined weights and this as discussed in part a is the IS estimator of $p_{tail}(\boldsymbol{a})$ too.
 
@@
 
-To calculate the RHS we use the Markov property and equation 8 and 9 in the coursework question sheet to get following equation:
+To calculate the RHS we can use equation 8 and 9 in the coursework question sheet:
 
@@
 \begin{aligned}
-Pr(\boldsymbol{X}=\boldsymbol{x}|\pi,\boldsymbol{P}_{\boldsymbol{T}}) = Pr(X_{1}=x_{1}|\pi,,\boldsymbol{P}_{\boldsymbol{T}})Pr(X_{2}=x_{2}|X_{1}=x_{1},,\boldsymbol{P}_{\boldsymbol{T}})\cdot \cdot \cdot Pr(X_{n}=x_{n}|X_{n-1}=x_{n-1},,\boldsymbol{P}_{\boldsymbol{T}}) = \\ \pi_{x_{1}}\prod_{i=2}^{n}p_{T}(X_{i}=x_{i}|X_{i-1}=x_{i-1})
+Pr(\boldsymbol{X}=\boldsymbol{x}|\boldsymbol{P}_{\boldsymbol{T}},\pi) = \pi_{x_{1}}\prod_{i=2}^{n}p_{T}(X_{i}=x_{i}|X_{i-1}=x_{i-1}) \hspace{1em} \text{with} \hspace{1em} x_{i} \in \{s,c,r\} \hspace{1em} \text{for} \hspace{1em} i \in \{1,..,n\}
 \end{aligned}
@@
 To calculate the ratio of the proposal density $q$ of the current and new values of $\boldsymbol{P}_{\boldsymbol{T}}$.
-Calculate the probability density of a Dirichlet($20 p^{cur}_{x}$) distribution (the proposal distribution) which I will denote $f(v)$ with $v$ equal to either $p^{cur}_{x}$ or $v= p^{new}_{x}$ for each of the rows of $\boldsymbol{P}_{\boldsymbol{T}}^{cur}$ or $\boldsymbol{P}_{\boldsymbol{T}}^{new}$ respectively.
+Calculate the probability density of a Dirichlet($20 p^{cur}_{x}$) distribution (the proposal distribution) which I will denote $f(v)$ with $v$ equal to either $p^{cur}_{x}$ or $p^{new}_{x}$ for each of the rows of $\boldsymbol{P}_{\boldsymbol{T}}^{cur}$ or $\boldsymbol{P}_{\boldsymbol{T}}^{new}$ respectively.
 Then calculate the the probability density of $\boldsymbol{P}_{\boldsymbol{T}}^{cur}$ and $\boldsymbol{P}_{\boldsymbol{T}}^{new}$ by finding the joint probability density of the rows $p_{x}$ by multiplying the probability densities of each row together (as independent can do this).
@@
 \begin{equation*}
- \frac{f(\boldsymbol{P}_{\boldsymbol{T}}^{cur})}{f(\boldsymbol{P}_{\boldsymbol{T}}^{new})} = \frac{\prod_{x=1}^{3}f(p_{x}^{cur})}{\prod_{x=1}^{3}f(p_{x}^{new})}  \hspace{1em} \text{with} \hspace{1em} x \in \{1=s,2=c,3=r\}
+ \frac{f(\boldsymbol{P}_{\boldsymbol{T}}^{cur})}{f(\boldsymbol{P}_{\boldsymbol{T}}^{new})} = \frac{\prod_{x=1}^{3}f(p_{x}^{cur})}{\prod_{x=1}^{3}f(p_{x}^{new})}  \hspace{1em} \text{with} \hspace{1em} \{1=s,2=c,3=r\}
 \end{equation*}

### 2022-12-30 Final.code.R

@@ -45,7 +45,7 @@ plot(x=x,y=target.dist(x),
      main="Plot 1: Comparing target and proposal pdf's",
      sub="Note: used a=2, b=3, s=1.25 with the green line seperating the regions of the graph")
 curve(gx.1(x),add=T,col="red")
-curve(gx.2,add=T,col="blue")
+curve(gx.2(x),add=T,col="blue")
 abline(v=-1.75,lty=2,col="green")
 abline(v=1.25,lty=2,col="green")
 text(x=-3,y=0.4,labels="Region 1")

### 2023-01-04 draft_report.Rmd
3rd review of report, slightly cleaning it up (one/two more small reviews needed)

@@
 For acceptance-rejection sampling (A-R sampling) we need to find a constant $M>0$ such that $\pi(x) \leq Mg(x), \forall x \in \mathbb{R}$.
-For $\pi(x)$ the target density (the density from which we want to generate a sample of random values) and $g(x)$ the proposal density defined in equations 1 and 2 above respectively.
+With $\pi(x)$ the target density (the density from which we want to generate a sample of random values) and $g(x)$ the proposal density defined in equations 1 and 2 above respectively.
 This constant M determines the acceptance probability which is $\frac{1}{M}$ and hence the efficiency of the algorithm.
@@
 To find M for each of the proposed $\lambda$ values I split the real domain of $x$ into 3 regions (see plot 1) that I call 1,2 and 3 which correspond to the different expressions of $\pi(x)$ depending on the value of $x$.
-Then study the relationship between $g(x)$ and $\pi(x)$ within these 3 regions to find $M_{1},M_{2},M_{3}$ which correspond to the smallest constant such that $\pi(x) \leq M_{i}g(x)$ for all possibles values of x in the ith region.
+Then study the relationship between $g(x)$ and $\pi(x)$ within these 3 regions to find $M_{1},M_{2},M_{3}$ which correspond to the smallest constant such that $\pi(x) \leq M_{i}g(x)$ for all possibles values of $x$ in the ith region.
 Then set $M=max(M_{1},M_{2},M_{3})$ as this is the smallest constant that satisfies $\pi(x) \leq Mg(x)$ for all possible values of $x \in \mathbb{R}$.
@@
 #### Region 1:
-For the first region ($-\infty, s-b$), $\pi(x)=0$ thus we can let $M_{1}$ be some arbitrarily small value greater than 0 say $\epsilon$.
+For the first region $x \in (-\infty, s-b)$, $\pi(x)=0$ thus we can let $M_{1}$ be some arbitrarily small value greater than 0 say $\epsilon$.
 This will satisfy $\pi(x) \leq M_{1}g(x)$ in the region specified as $g(x) > 0$ for all values of $x$ in this region.
@@
 #### Region 2:
-For the second region [$s-b,s$], $\pi(x)=\frac{1}{2b}$.
+For the second region $x \in [s-b,s]$, $\pi(x)=\frac{1}{2b}$.
 Considering that $g(x)$ is an increasing function in this region (see plot 1) it will be at it's minimum value at $x=s-b$ and the value of $\pi(x)$ is constant.
@@
 #### Region 3:
-For the third region ($s,\infty$), $\pi(x)=\frac{as^{a}}{2x^{a+1}}$.
+For the third region $x \in (s,\infty)$, $\pi(x)=\frac{as^{a}}{2x^{a+1}}$.
 I note that $\pi(x)$ is a decreasing function in this region (see plot 1) and initially $g(x)$ is a smaller but increasing function (see plot 1), until it reaches its maximum value at which point it is larger than $\pi(x)$.
@@
 #### Sampling from the proposal distribution:
-To sample from $g(x)$, use the inverse transformation technique by drawing $u$ from a Uniform(0,1) distribution then if $u<0.5$ set $x=b+\frac{log(2u)}{\lambda}$, else set $x=b-\frac{2-2u}{\lambda}$.
-To obtain the value $x$, drawn from $g(x)$ the proposal distribution.
+To sample from the proposal distribution, use the inverse transformation technique by drawing $u$ from a Uniform(0,1) distribution then if $u<0.5$ set $x=b+\frac{log(2u)}{\lambda}$, else set $x=b-\frac{2-2u}{\lambda}$.
+To obtain the value $x$, drawn from the proposal distribution.
 
@@
 
-* $s-b <x$, check if $y \leq \frac{\pi(x)}{Mg(x)}=\frac{1}{2b}\frac{2}{M\lambda exp(-\lambda|x-b|)}$
+* $s-b \leq x \leq s$, check if $y \leq \frac{\pi(x)}{Mg(x)}=\frac{1}{2b}\frac{2}{M\lambda exp(-\lambda|x-b|)}$
 
-If either of these checks are true then $x$ is a draw from the target distribution $\pi(x)$ and the value is saved, if these checks are not true or $x<s-b$ then $x$ is not a draw from $\pi(x)$ and the value of $x$ is disregarded.
-Repeat this process until the desired number of samples from $\pi(x)$ has been attained.
+If either of these checks are true then $x$ is a draw from the target distribution with density $\pi(x)$ and the value is saved, if these checks are not true or $x<s-b$ then $x$ is not a draw from $\pi(x)$ and the value of $x$ is disregarded.
+Repeat this process until the desired number of samples from the target distribution has been attained.
 
@@
 \begin{equation*}
-Pr(U>\boldsymbol{b})=\prod_{i=1}^{d}Pr\left(Z>\frac{b_{i}-u_{i-1}}{\sqrt{t_{i}-t_{i-1}}}\right)=\prod_{i=1}^{d}\int_{s_{i}}^{\infty}\phi(x)dx=\prod_{i=1}^{d}\int_{-\infty}^{\infty}\frac{\phi(x)}{q_{i}(x)}q_{i}(x)I(x>s_{i})dx = E\left[\prod_{i=1}^{d} \frac{\phi(V_{i})}{q_{i}(V_{i})}I(V_{i}>s_{i})\right]
+Pr(U>\boldsymbol{b})=\prod_{i=1}^{d}Pr\left(Z>\frac{b_{i}-u_{i-1}}{\sqrt{t_{i}-t_{i-1}}}\right)=\prod_{i=1}^{d}\int_{s_{i}}^{\infty}\phi(x)dx=\prod_{i=1}^{d}\int_{-\infty}^{\infty}\frac{\phi(x)}{q_{i}(x)}q_{i}(x)I(x>s_{i})dx = E\left[\prod_{i=1}^{d} w_{i}(V_{i})I(V_{i}>s_{i})\right]
 \end{equation*}
@@
 With $V_{i}, i \in \{1,..,d\}$ i.i.d.
-random variables from the proposal distribution, $q_{i}$ the density of the proposal distribution, $\phi$ the density of a standard normal distribution and $I$ the indicator function.
+random variables from the proposal distribution, $q_{i}$ the density of the proposal distribution, $\phi$ the density of a standard normal distribution, $w_{i}(V_{i})$ the (unnormalized) importance weights (equal to $\frac{\phi(V_{i})}{q_{i}(V_{i})}$) and $I$ the indicator function.
 
@@
 
-Where $w_{i}^{(j)}$ is $\frac{\phi(V_{i})}{q_{i}(V_{i})}$ on the jth iteration and the superscript $(j)$ denoting the jth sampling.
+Where $w_{i}^{(j)}$ is $\frac{\phi(V_{i})}{q_{i}(V_{i})}$ the superscript $(j)$ denotes the jth sampling iteration.
 Therefore, to not waste any sample values let the proposal distributions be exponential distributions shifted by $s_{i}$ as all sample values (denoted $v_{i}$) drawn from these proposal distributions will satisfy $v_{i}>s_{i}$ for $i \in \{1,...,d\}$.
@@
 #### Drawing values from proposal distributions:
-Next for each iteration j we must simulate the random walk $U$ using equation (3), the initial value $u_{0}$ and times $t_{i}, i \in \{0,..,d\}$ and storing the values of the random walk $\boldsymbol{u}$ as mentioned in part a.
-Using these values and the relevant proposal distribution of an exponential distribution shifted by $s_{i}$ (described in part a) to generate $v_{i}, i \in \{1,..,d\}$.
+Next for each iteration j we must simulate the random walk $U$ using equation (3), the initial value $u_{0}$ and times $t_{i}, i \in \{0,..,d\}$ and storing the values of the generated random walk $\boldsymbol{u}$ as mentioned in part a.
+Using these values and the relevant proposal distribution of an exponential distribution shifted by $s_{i}$ (described in part a) then generate $v_{i}, i \in \{1,..,d\}$.
 
@@
 
-Then calculate the product of the weights $\prod_{i=1}^{n}w_{i}$ and save this as the jth value in a vector.
+Then calculate the product of the weights for this jth iteration $\hat{w}^{(j)}:= \prod_{i=1}^{n}w_{i}^{(j)}$ and save this value in a vector in element j.
 
 #### Calculating probability:
-To calculate $p_{tail}(\boldsymbol{b})$, the previous 2 steps are performed n times thus we have n values of the product of the weights (which I will denote $\hat{w}^{(j)}$ where the value of $j$ denotes the jth iteration) which we then use to calculate the importance sampling (IS) estimator using equation (4) in part a:
+To calculate $p_{tail}(\boldsymbol{a})$, the previous 2 steps are performed n times thus we have n values of the product of the weights which we then use to calculate the importance sampling (IS) estimator using equation (4) in part a:
 
@@
 #### Drawing new proposal value:
-For each iteration j start by drawing $\boldsymbol{P}_{\boldsymbol{T}}^{new}$ the proposed value of $\boldsymbol{P}_{\boldsymbol{T}}$ from the proposal density by drawing each $p_{x}'$ with $x \in \{s,c,r\}$ from a Dirichlet $(20p_{x}^{(cur)})$ distribution and combining these to get $\boldsymbol{P}_{\boldsymbol{T}}^{new}=(p_{s}',p_{c}',p_{r}')^{T}$.
+For each iteration j start by generating $\boldsymbol{P}_{\boldsymbol{T}}^{new}$ the proposed value of $\boldsymbol{P}_{\boldsymbol{T}}$ from the proposal density by drawing each $p_{x}'$ with $x \in \{s,c,r\}$ from a Dirichlet $(20p_{x}^{(cur)})$ distribution and combining these to get $\boldsymbol{P}_{\boldsymbol{T}}^{new}=(p_{s}',p_{c}',p_{r}')^{T}$.
 
@@
 #### Accept or reject proposal:
-Calculate $\alpha = min \left( 1,\frac{p(\boldsymbol{P}_{\boldsymbol{T}}^{new}|X,\pi)}{p(\boldsymbol{P}_{\boldsymbol{T}}^{cur}|X,\pi)}\frac{f(\boldsymbol{P}_{\boldsymbol{T}}^{cur})}{f(\boldsymbol{P}_{\boldsymbol{T}}^{new})}\right)$ using the ratios calculated above and generate $u$ a value drawn from a Uniform(0,1) distribution and if $u<=\alpha$ set $\boldsymbol{P}_{\boldsymbol{T}}^{new}$ to be the jth value of $\boldsymbol{P}_{\boldsymbol{T}}$ and update $\boldsymbol{P}_{\boldsymbol{T}}^{cur}$ to be the value of $\boldsymbol{P}_{\boldsymbol{T}}^{new}$.
-Otherwise, set $\boldsymbol{P}_{\boldsymbol{T}}^{cur}$ to be jth value of $\boldsymbol{P}_{\boldsymbol{T}}$ and don't update the value of $\boldsymbol{P}_{\boldsymbol{T}}^{cur}$.
+Calculate $\alpha = min \left( 1,\frac{p(\boldsymbol{P}_{\boldsymbol{T}}^{new}|X,\pi)}{p(\boldsymbol{P}_{\boldsymbol{T}}^{cur}|X,\pi)}\frac{f(\boldsymbol{P}_{\boldsymbol{T}}^{cur})}{f(\boldsymbol{P}_{\boldsymbol{T}}^{new})}\right)$ using the ratios calculated above and generate $u$ a value drawn from a Uniform(0,1) distribution and if $u<=\alpha$ set $\boldsymbol{P}_{\boldsymbol{T}}^{new}$ to be the jth sample value of $\boldsymbol{P}_{\boldsymbol{T}}$ and update $\boldsymbol{P}_{\boldsymbol{T}}^{cur}$ to be the value of $\boldsymbol{P}_{\boldsymbol{T}}^{new}$.
+Otherwise, set $\boldsymbol{P}_{\boldsymbol{T}}^{cur}$ to be jth sample value of $\boldsymbol{P}_{\boldsymbol{T}}$ and don't update the value of $\boldsymbol{P}_{\boldsymbol{T}}^{cur}$.
 
 #### Iterate:
-Repeat until we have attained n+b+1 sample values of $\boldsymbol{P}_{\boldsymbol{T}}$ where n is the number of desired samples and b is the specified burn in number.
+Repeat until we have attained n+b+1 sample values of $\boldsymbol{P}_{\boldsymbol{T}}$ where n is the number of desired posterior sample values and b is the specified burn in number.
 Then throw away the first b+1 values of $\boldsymbol{P}_{\boldsymbol{T}}$.

### 2023-01-04 Final.code.R
Chekcing Q4 code and updating bi and iii for list return

@@ -305,25 +305,23 @@ linRegMixEstep <- function(piCur,aCur,bCur,sCur,x,Y){
 ### Part b ###
 
 ### (i)
-
 calcNewCoefs <- function(W,x,Y){
   #Setting up matrix to store values of new coefficients in
-  new.coef <- matrix(data=NA,ncol = 2,nrow=dim(W)[2])
+  new.coef <- matrix(data=NA,ncol = dim(W)[2],nrow=2)
   #Creating design matrix
   X <- matrix(c(rep(1,length(x)),x),ncol=2)
   
-  #Calculating a^{new) and b^{new}
+  #Calculating a^{new) and b^{new} using equation 19 in coursework sheet
   for (k in 1:dim(W)[2]){
-    new.coef[k,] <- solve(t(X)%*%diag(W[,k])%*%X)%*%t(X)%*%diag(W[,k])%*%Y
+    new.coef[,k] <- solve(t(X)%*%diag(W[,k])%*%X)%*%t(X)%*%diag(W[,k])%*%Y
   }
-  return(t(new.coef))
+  return(list("a.new"=new.coef[1,],"b.new"=new.coef[2,]))
 }
 
 ### (ii)
-
 calcNewSd <- function(W,x,Y,a,b){
   new.sd <- NULL
-  #calculating new sd
+  #calculating new sd using equation 20 in coursework sheet
   for (k in 1:dim(W)[2]){
     new.sd[k] <- sqrt(sum(W[,k]*(Y-a[k]-b[k]*x)^2)/sum(W[,k]))
   }
@@ -336,15 +334,15 @@ linRegMixMstep <- function(W,x,Y){
   #calculating new coefficents for a and b
   new.coef <- calcNewCoefs(W,x,Y)
   
-  #calculating new sd
-  new.sd <- calcNewSd(W,x,Y,new.coef[1,],new.coef[2,]) 
+  #calculating new sd vector
+  new.sd <- calcNewSd(W,x,Y,new.coef$a.new,new.coef$b.new) 
   
-  #Calculating new pi
+  #Calculating new pi using equation 18 in coursework sheet
   new.pi <- NULL
   for (k in 1:dim(W)[2]){
     new.pi[k] <- sum(W[,k])/dim(W)[1]
   }
-  return(list(new.pi,new.coef[1,],new.coef[2,],new.sd))
[…]

### 2023-01-04 Final.code.R
Going over code to see that functions and inputs correct

@@ -192,7 +192,7 @@ round(is.estimate,4)
 calcDirichletPDF <- function(x,a){
   #Calculating log density of x under Dirichlet distribution with parameters specified by a
   log.density <- lgamma(sum(a))-sum(lgamma(a))+sum((a-1)*log(x))
-  return(c(log.density))
+  return(log.density)
 }
 
 ### Part c ###

### 2023-01-07 draft_report.Rmd
Final review of report

@@
 For multiple possible values of $\lambda$ in $g(x)$ the value of $\lambda$ that results in a smaller constant M will have a higher acceptance probability and thus will be our preferred value of $\lambda$ for the algorithm.
-To find M for each of the proposed $\lambda$ values I split the real domain of $x$ into 3 regions (see plot 1) that I call 1,2 and 3 which correspond to the different expressions of $\pi(x)$ depending on the value of $x$.
+To find M for each of the proposed $\lambda$ values I split the real domain of $x$ into 3 regions (see plot 1) that I call 1,2 and 3 which correspond to the different forms of the equation of $\pi(x)$ depending on the value of $x$.
 Then study the relationship between $g(x)$ and $\pi(x)$ within these 3 regions to find $M_{1},M_{2},M_{3}$ which correspond to the smallest constant such that $\pi(x) \leq M_{i}g(x)$ for all possibles values of $x$ in the ith region.
@@
 ## Part b:
-To implement A-R sampling using the proposal density with $\lambda=0.5$ and thus $M=3.2e^{0.875}$ to obtain samples from $\pi(x)$ with parameters $s$, $a$ and $b$ I use the following algorithm:
+To implement A-R sampling using the proposal density with $\lambda=0.5$ and thus $M=3.2e^{0.875}$ to obtain samples from $\pi(x)$ with parameters $s$, $a$ and $b$ I used the following algorithm:
 
